\documentclass[11pt,letterpaper,oneside]{memoir}
\usepackage{charter}

% Set the typeblock size to ensure about 65 characters per line (with Charter 11pt, that's 
% 26 picas).
\settypeblocksize{8.5in}{26pc}{*}
% assuming room for a 3-hole punch (0.625in on the spine)
%\setlrmargins{1.625in}{*}{*}
\setlrmargins{12.5pc}{*}{*}
\setulmargins{1.25in}{*}{*}
\checkandfixthelayout

\pagestyle{companion}

%% Bringhurst chapter style
\makechapterstyle{bringhurst}{%
\renewcommand{\chapterheadstart}{}
\renewcommand{\printchaptername}{}
\renewcommand{\chapternamenum}{}
\renewcommand{\printchapternum}{}
\renewcommand{\afterchapternum}{}
\renewcommand{\printchaptertitle}[1]{%
\raggedright\Large\scshape\MakeLowercase{##1}}
\renewcommand{\afterchaptertitle}{%
\vskip\onelineskip \hrule\vskip\onelineskip}
}

\setsecheadstyle{\raggedright\scshape\MakeLowercase}
\setbeforesecskip{-\onelineskip}
\setaftersecskip{\onelineskip}

\chapterstyle{bringhurst}

\begin{document}
\chapter[2 John]{The Second Epistle of John}

The elder unto the elect lady and her children, whom I love in the
truth; and not I only, but also all they that have known the truth;
For the truth's sake, which dwelleth in us, and shall be with us for
ever.  Grace be with you, mercy, \emph{and} peace, from God the
Father, and from the Lord Jesus Christ, the Son of the Father, in
truth and love.

I rejoiced greatly that I found of thy children walking in truth, as
we have received a commandment from the Father.  And now I beseech
thee, lady, not as though I wrote a new commandment unto thee, but
that which we had from the beginning, that we love one another.  And
this is love, that we walk after his commandments. This is the
commandment, That, as ye have heard from the beginning, ye should walk
in it.  For many deceivers are entered into the world, who confess not
that Jesus Christ is come in the flesh. This is a deceiver and an
antichrist.  Look to yourselves, that we lose not those things which
we have wrought, but that we receive a full reward.  Whosoever
transgresseth, and abideth not in the doctrine of Christ, hath not
God. He that abideth in the doctrine of Christ, he hath both the
Father and the Son.  If there come any unto you, and bring not this
doctrine, receive him not into \emph{your} house, neither bid him God
speed:  For he that biddeth him God speed is partaker of his evil
deeds.

Having many things to write unto you, I would not \emph{write} with
paper and ink: but I trust to come unto you, and speak face to face,
that our joy may be full.  The children of thy elect sister greet
thee. Amen.

\end{document}