\documentclass[11pt,letterpaper,oneside]{memoir}
\usepackage{charter}

% Set the typeblock size to ensure about 65 characters per line (with Charter 11pt, that's 
% 26 picas).
\settypeblocksize{8.5in}{26pc}{*}
% assuming room for a 3-hole punch (0.625in on the spine)
%\setlrmargins{1.625in}{*}{*}
\setlrmargins{12.5pc}{*}{*}
\setulmargins{1.25in}{*}{*}
\checkandfixthelayout

\pagestyle{companion}

%% Bringhurst chapter style
\makechapterstyle{bringhurst}{%
\renewcommand{\chapterheadstart}{}
\renewcommand{\printchaptername}{}
\renewcommand{\chapternamenum}{}
\renewcommand{\printchapternum}{}
\renewcommand{\afterchapternum}{}
\renewcommand{\printchaptertitle}[1]{%
\raggedright\Large\scshape\MakeLowercase{##1}}
\renewcommand{\afterchaptertitle}{%
\vskip\onelineskip \hrule\vskip\onelineskip}
}

\setsecheadstyle{\raggedright\scshape\MakeLowercase}
\setbeforesecskip{-\onelineskip}
\setaftersecskip{\onelineskip}

\chapterstyle{bringhurst}

\begin{document}
\chapter[3 John]{The Third Epistle of John}

The elder unto the wellbeloved Gaius, whom I love in the truth.

Beloved, I wish above all things that thou mayest prosper and be in
health, even as thy soul prospereth.  For I rejoiced greatly, when the
brethren came and testified of the truth that is in thee, even as thou
walkest in the truth.  I have no greater joy than to hear that my
children walk in truth.  Beloved, thou doest faithfully whatsoever
thou doest to the brethren, and to strangers;  Which have borne
witness of thy charity before the church: whom if thou bring forward
on their journey after a godly sort, thou shalt do well:  Because that
for his name's sake they went forth, taking nothing of the Gentiles.
We therefore ought to receive such, that we might be fellowhelpers to
the truth.

I wrote unto the church: but Diotrephes, who loveth to have the
preeminence among them, receiveth us not.  Wherefore, if I come, I
will remember his deeds which he doeth, prating against us with
malicious words: and not content therewith, neither doth he himself
receive the brethren, and forbiddeth them that would, and casteth
\emph{them} out of the church.

Beloved, follow not that which is evil, but that which is good. He
that doeth good is of God: but he that doeth evil hath not seen God.
Demetrius hath good report of all \emph{men}, and of the truth itself:
yea, and we \emph{also} bear record; and ye know that our record is
true.

I had many things to write, but I will not with ink and pen write
unto thee:  But I trust I shall shortly see thee, and we shall speak
face to face. Peace \emph{be} to thee. \emph{Our} friends salute thee.
Greet the friends by name.

\end{document}