\documentclass[11pt,letterpaper,oneside]{memoir}
\usepackage{charter}

% Set the typeblock size to ensure about 65 characters per line (with Charter 11pt, that's 
% 26 picas).
\settypeblocksize{8.5in}{26pc}{*}
% assuming room for a 3-hole punch (0.625in on the spine)
%\setlrmargins{1.625in}{*}{*}
\setlrmargins{12.5pc}{*}{*}
\setulmargins{1.25in}{*}{*}
\checkandfixthelayout

\pagestyle{companion}

%% Bringhurst chapter style
\makechapterstyle{bringhurst}{%
\renewcommand{\chapterheadstart}{}
\renewcommand{\printchaptername}{}
\renewcommand{\chapternamenum}{}
\renewcommand{\printchapternum}{}
\renewcommand{\afterchapternum}{}
\renewcommand{\printchaptertitle}[1]{%
\raggedright\Large\scshape\MakeLowercase{##1}}
\renewcommand{\afterchaptertitle}{%
\vskip\onelineskip \hrule\vskip\onelineskip}
}

\setsecheadstyle{\raggedright\scshape\MakeLowercase}
\setbeforesecskip{-\onelineskip}
\setaftersecskip{\onelineskip}

\chapterstyle{bringhurst}

\begin{document}
\chapter[Exodus]{The Second Book of Moses, called Exodus}
Now these \emph{are} the names of the children of Israel, which came
into Egypt; every man and his household came with Jacob. Reuben, Simeon,
Levi, and Judah, Issachar, Zebulun, and Benjamin, Dan, and Naphtali,
Gad, and Asher. And all the souls that came out of the loins of Jacob
were seventy souls: for Joseph was in Egypt \emph{already. }And Joseph
died, and all his brethren, and all that generation.

And the children of Israel were fruitful, and increased abundantly, and
multiplied, and waxed exceeding mighty; and the land was filled with
them. Now there arose up a new king over Egypt, which knew not Joseph.
And he said unto his people, Behold, the people of the children of
Israel \emph{are} more and mightier than we: Come on, let us deal wisely
with them; lest they multiply, and it come to pass, that, when there
falleth out any war, they join also unto our enemies, and fight against
us, and \emph{so} get them up out of the land. Therefore they did set
over them taskmasters to afflict them with their burdens. And they built
for Pharaoh treasure cities, Pithom and Raamses. But the more they
afflicted them, the more they multiplied and grew. And they were grieved
because of the children of Israel. And the Egyptians made the children
of Israel to serve with rigour: And they made their lives bitter with
hard bondage, in morter, and in brick, and in all manner of service in
the field: all their service, wherein they made them serve, \emph{was}
with rigour.

And the king of Egypt spake to the Hebrew midwives, of which the name of
the one \emph{was} Shiphrah, and the name of the other Puah: And he
said, When ye do the office of a midwife to the Hebrew women, and see
\emph{them} upon the stools; if it \emph{be} a son, then ye shall kill
him: but if it \emph{be} a daughter, then she shall live. But the
midwives feared God, and did not as the king of Egypt commanded them,
but saved the men children alive. And the king of Egypt called for the
midwives, and said unto them, Why have ye done this thing, and have
saved the men children alive? And the midwives said unto Pharaoh,
Because the Hebrew women \emph{are} not as the Egyptian women; for they
\emph{are} lively, and are delivered ere the midwives come in unto them.
Therefore God dealt well with the midwives: and the people multiplied,
and waxed very mighty. And it came to pass, because the midwives feared
God, that he made them houses. And Pharaoh charged all his people,
saying, Every son that is born ye shall cast into the river, and every
daughter ye shall save alive.

And there went a man of the house of Levi, and took \emph{to wife} a
daughter of Levi. And the woman conceived, and bare a son: and when she
saw him that he \emph{was a} goodly \emph{child,} she hid him three
months. And when she could not longer hide him, she took for him an ark
of bulrushes, and daubed it with slime and with pitch, and put the child
therein; and she laid \emph{it} in the flags by the river's brink. And
his sister stood afar off, to wit what would be done to him.

And the daughter of Pharaoh came down to wash \emph{herself} at the
river; and her maidens walked along by the river's side; and when she
saw the ark among the flags, she sent her maid to fetch it. And when she
had opened \emph{it,} she saw the child: and, behold, the babe wept. And
she had compassion on him, and said, This \emph{is one} of the Hebrews'
children. Then said his sister to Pharaoh's daughter, Shall I go and
call to thee a nurse of the Hebrew women, that she may nurse the child
for thee? And Pharaoh's daughter said to her, Go. And the maid went and
called the child's mother. And Pharaoh's daughter said unto her, Take
this child away, and nurse it for me, and I will give \emph{thee} thy
wages. And the woman took the child, and nursed it. And the child grew,
and she brought him unto Pharaoh's daughter, and he became her son. And
she called his name Moses: and she said, Because I drew him out of the
water.

And it came to pass in those days, when Moses was grown, that he went
out unto his brethren, and looked on their burdens: and he spied an
Egyptian smiting an Hebrew, one of his brethren. And he looked this way
and that way, and when he saw that \emph{there was} no man, he slew the
Egyptian, and hid him in the sand. And when he went out the second day,
behold, two men of the Hebrews strove together: and he said to him that
did the wrong, Wherefore smitest thou thy fellow? And he said, Who made
thee a prince and a judge over us? intendest thou to kill me, as thou
killedst the Egyptian? And Moses feared, and said, Surely this thing is
known. Now when Pharaoh heard this thing, he sought to slay Moses. But
Moses fled from the face of Pharaoh, and dwelt in the land of Midian:
and he sat down by a well. Now the priest of Midian had seven daughters:
and they came and drew \emph{water,} and filled the troughs to water
their father's flock. And the shepherds came and drove them away: but
Moses stood up and helped them, and watered their flock. And when they
came to Reuel their father, he said, How \emph{is it that} ye are come
so soon to day? And they said, An Egyptian delivered us out of the hand
of the shepherds, and also drew \emph{water} enough for us, and watered
the flock. And he said unto his daughters, And where \emph{is} he? why
\emph{is} it \emph{that} ye have left the man? call him, that he may eat
bread. And Moses was content to dwell with the man: and he gave Moses
Zipporah his daughter. And she bare \emph{him} a son, and he called his
name Gershom: for he said, I have been a stranger in a strange land.

And it came to pass in process of time, that the king of Egypt died: and
the children of Israel sighed by reason of the bondage, and they cried,
and their cry came up unto God by reason of the bondage. And God heard
their groaning, and God remembered his covenant with Abraham, with
Isaac, and with Jacob. And God looked upon the children of Israel, and
God had respect unto \emph{them.}

Now Moses kept the flock of Jethro his father in law, the priest of
Midian: and he led the flock to the backside of the desert, and came to
the mountain of God, \emph{even} to Horeb. And the angel of the
\textsc{Lord} appeared unto him in a flame of fire out of the midst of a
bush: and he looked, and, behold, the bush burned with fire, and the
bush \emph{was} not consumed. And Moses said, I will now turn aside, and
see this great sight, why the bush is not burnt. And when the
\textsc{Lord} saw that he turned aside to see, God called unto him out
of the midst of the bush, and said, Moses, Moses. And he said, Here
\emph{am} I. And he said, Draw not nigh hither: put off thy shoes from
off thy feet, for the place whereon thou standest is holy ground.
Moreover he said, I \emph{am} the God of thy father, the God of Abraham,
the God of Isaac, and the God of Jacob. And Moses hid his face; for he
was afraid to look upon God.

And the \textsc{Lord} said, I have surely seen the affliction of my
people which \emph{are} in Egypt, and have heard their cry by reason of
their taskmasters; for I know their sorrows; And I am come down to
deliver them out of the hand of the Egyptians, and to bring them up out
of that land unto a good land and a large, unto a land flowing with milk
and honey; unto the place of the Canaanites, and the Hittites, and the
Amorites, and the Perizzites, and the Hivites, and the Jebusites. Now
therefore, behold, the cry of the children of Israel is come unto me:
and I have also seen the oppression wherewith the Egyptians oppress
them. Come now therefore, and I will send thee unto Pharaoh, that thou
mayest bring forth my people the children of Israel out of Egypt.

And Moses said unto God, Who \emph{am} I, that I should go unto Pharaoh,
and that I should bring forth the children of Israel out of Egypt? And
he said, Certainly I will be with thee; and this \emph{shall be} a token
unto thee, that I have sent thee: When thou hast brought forth the
people out of Egypt, ye shall serve God upon this mountain. And Moses
said unto God, Behold, \emph{when} I come unto the children of Israel,
and shall say unto them, The God of your fathers hath sent me unto you;
and they shall say to me, What is his name? what shall I say unto them?
And God said unto Moses, I AM THAT I AM: and he said, Thus shalt thou
say unto the children of Israel, I AM hath sent me unto you. And God
said moreover unto Moses, Thus shalt thou say unto the children of
Israel, The \textsc{Lord} God of your fathers, the God of Abraham, the
God of Isaac, and the God of Jacob, hath sent me unto you: this
\emph{is} my name for ever, and this \emph{is} my memorial unto all
generations. Go, and gather the elders of Israel together, and say unto
them, The \textsc{Lord} God of your fathers, the God of Abraham, of
Isaac, and of Jacob, appeared unto me, saying, I have surely visited
you, and \emph{seen} that which is done to you in Egypt: And I have
said, I will bring you up out of the affliction of Egypt unto the land
of the Canaanites, and the Hittites, and the Amorites, and the
Perizzites, and the Hivites, and the Jebusites, unto a land flowing with
milk and honey. And they shall hearken to thy voice: and thou shalt
come, thou and the elders of Israel, unto the king of Egypt, and ye
shall say unto him, The \textsc{Lord} God of the Hebrews hath met with
us: and now let us go, we beseech thee, three days' journey into the
wilderness, that we may sacrifice to the \textsc{Lord} our God.

And I am sure that the king of Egypt will not let you go, no, not by a
mighty hand. And I will stretch out my hand, and smite Egypt with all my
wonders which I will do in the midst thereof: and after that he will let
you go. And I will give this people favour in the sight of the
Egyptians: and it shall come to pass, that, when ye go, ye shall not go
empty: But every woman shall borrow of her neighbour, and of her that
sojourneth in her house, jewels of silver, and jewels of gold, and
raiment: and ye shall put \emph{them} upon your sons, and upon your
daughters; and ye shall spoil the Egyptians.

And Moses answered and said, But, behold, they will not believe me, nor
hearken unto my voice: for they will say, The \textsc{Lord} hath not
appeared unto thee. And the \textsc{Lord} said unto him, What \emph{is}
that in thine hand? And he said, A rod. And he said, Cast it on the
ground. And he cast it on the ground, and it became a serpent; and Moses
fled from before it. And the \textsc{Lord} said unto Moses, Put forth
thine hand, and take it by the tail. And he put forth his hand, and
caught it, and it became a rod in his hand: That they may believe that
the \textsc{Lord} God of their fathers, the God of Abraham, the God of
Isaac, and the God of Jacob, hath appeared unto thee.

And the \textsc{Lord} said furthermore unto him, Put now thine hand into
thy bosom. And he put his hand into his bosom: and when he took it out,
behold, his hand \emph{was} leprous as snow. And he said, Put thine hand
into thy bosom again. And he put his hand into his bosom again; and
plucked it out of his bosom, and, behold, it was turned again as his
\emph{other} flesh. And it shall come to pass, if they will not believe
thee, neither hearken to the voice of the first sign, that they will
believe the voice of the latter sign. And it shall come to pass, if they
will not believe also these two signs, neither hearken unto thy voice,
that thou shalt take of the water of the river, and pour \emph{it} upon
the dry \emph{land:} and the water which thou takest out of the river
shall become blood upon the dry \emph{land.}

And Moses said unto the \textsc{Lord}, O my Lord, I \emph{am} not
eloquent, neither heretofore, nor since thou hast spoken unto thy
servant: but I \emph{am} slow of speech, and of a slow tongue. And the
\textsc{Lord} said unto him, Who hath made man's mouth? or who maketh
the dumb, or deaf, or the seeing, or the blind? have not I the
\textsc{Lord}? Now therefore go, and I will be with thy mouth, and teach
thee what thou shalt say. And he said, O my Lord, send, I pray thee, by
the hand \emph{of him whom} thou wilt send. And the anger of the
\textsc{Lord} was kindled against Moses, and he said, \emph{Is} not
Aaron the Levite thy brother? I know that he can speak well. And also,
behold, he cometh forth to meet thee: and when he seeth thee, he will be
glad in his heart. And thou shalt speak unto him, and put words in his
mouth: and I will be with thy mouth, and with his mouth, and will teach
you what ye shall do. And he shall be thy spokesman unto the people: and
he shall be, \emph{even} he shall be to thee instead of a mouth, and
thou shalt be to him instead of God. And thou shalt take this rod in
thine hand, wherewith thou shalt do signs.

And Moses went and returned to Jethro his father in law, and said unto
him, Let me go, I pray thee, and return unto my brethren which
\emph{are} in Egypt, and see whether they be yet alive. And Jethro said
to Moses, Go in peace. And the \textsc{Lord} said unto Moses in Midian,
Go, return into Egypt: for all the men are dead which sought thy life.
And Moses took his wife and his sons, and set them upon an ass, and he
returned to the land of Egypt: and Moses took the rod of God in his
hand. And the \textsc{Lord} said unto Moses, When thou goest to return
into Egypt, see that thou do all those wonders before Pharaoh, which I
have put in thine hand: but I will harden his heart, that he shall not
let the people go. And thou shalt say unto Pharaoh, Thus saith the
\textsc{Lord}, Israel \emph{is} my son, \emph{even} my firstborn: And I
say unto thee, Let my son go, that he may serve me: and if thou refuse
to let him go, behold, I will slay thy son, \emph{even} thy firstborn.

And it came to pass by the way in the inn, that the \textsc{Lord} met
him, and sought to kill him. Then Zipporah took a sharp stone, and cut
off the foreskin of her son, and cast \emph{it} at his feet, and said,
Surely a bloody husband \emph{art} thou to me. So he let him go: then
she said, A bloody husband \emph{thou art,} because of the circumcision.

And the \textsc{Lord} said to Aaron, Go into the wilderness to meet
Moses. And he went, and met him in the mount of God, and kissed him. And
Moses told Aaron all the words of the \textsc{Lord} who had sent him,
and all the signs which he had commanded him.

And Moses and Aaron went and gathered together all the elders of the
children of Israel: And Aaron spake all the words which the
\textsc{Lord} had spoken unto Moses, and did the signs in the sight of
the people. And the people believed: and when they heard that the
\textsc{Lord} had visited the children of Israel, and that he had looked
upon their affliction, then they bowed their heads and worshipped.

And afterward Moses and Aaron went in, and told Pharaoh, Thus saith the
\textsc{Lord} God of Israel, Let my people go, that they may hold a
feast unto me in the wilderness. And Pharaoh said, Who \emph{is} the
\textsc{Lord}, that I should obey his voice to let Israel go? I know not
the \textsc{Lord}, neither will I let Israel go. And they said, The God
of the Hebrews hath met with us: let us go, we pray thee, three days'
journey into the desert, and sacrifice unto the \textsc{Lord} our God;
lest he fall upon us with pestilence, or with the sword. And the king of
Egypt said unto them, Wherefore do ye, Moses and Aaron, let the people
from their works? get you unto your burdens. And Pharaoh said, Behold,
the people of the land now \emph{are} many, and ye make them rest from
their burdens. And Pharaoh commanded the same day the taskmasters of the
people, and their officers, saying, Ye shall no more give the people
straw to make brick, as heretofore: let them go and gather straw for
themselves. And the tale of the bricks, which they did make heretofore,
ye shall lay upon them; ye shall not diminish \emph{ought} thereof: for
they \emph{be} idle; therefore they cry, saying, Let us go \emph{and}
sacrifice to our God. Let there more work be laid upon the men, that
they may labour therein; and let them not regard vain words.

And the taskmasters of the people went out, and their officers, and they
spake to the people, saying, Thus saith Pharaoh, I will not give you
straw. Go ye, get you straw where ye can find it: yet not ought of your
work shall be diminished. So the people were scattered abroad throughout
all the land of Egypt to gather stubble instead of straw. And the
taskmasters hasted \emph{them,} saying, Fulfil your works, \emph{your}
daily tasks, as when there was straw. And the officers of the children
of Israel, which Pharaoh's taskmasters had set over them, were beaten,
\emph{and} demanded, Wherefore have ye not fulfilled your task in making
brick both yesterday and to day, as heretofore?

Then the officers of the children of Israel came and cried unto Pharaoh,
saying, Wherefore dealest thou thus with thy servants? There is no straw
given unto thy servants, and they say to us, Make brick: and, behold,
thy servants \emph{are} beaten; but the fault \emph{is} in thine own
people. But he said, Ye \emph{are} idle, \emph{ye are} idle: therefore
ye say, Let us go \emph{and} do sacrifice to the \textsc{Lord}. Go
therefore now, \emph{and} work; for there shall no straw be given you,
yet shall ye deliver the tale of bricks. And the officers of the
children of Israel did see \emph{that} they \emph{were} in evil
\emph{case,} after it was said, Ye shall not minish \emph{ought} from
your bricks of your daily task.

And they met Moses and Aaron, who stood in the way, as they came forth
from Pharaoh: And they said unto them, The \textsc{Lord} look upon you,
and judge; because ye have made our savour to be abhorred in the eyes of
Pharaoh, and in the eyes of his servants, to put a sword in their hand
to slay us. And Moses returned unto the \textsc{Lord}, and said, Lord,
wherefore hast thou \emph{so} evil entreated this people? why \emph{is}
it \emph{that} thou hast sent me? For since I came to Pharaoh to speak
in thy name, he hath done evil to this people; neither hast thou
delivered thy people at all.

Then the \textsc{Lord} said unto Moses, Now shalt thou see what I will
do to Pharaoh: for with a strong hand shall he let them go, and with a
strong hand shall he drive them out of his land. And God spake unto
Moses, and said unto him, I \emph{am} the \textsc{Lord}: And I appeared
unto Abraham, unto Isaac, and unto Jacob, by \emph{the name of} God
Almighty, but by my name JEHOVAH was I not known to them. And I have
also established my covenant with them, to give them the land of Canaan,
the land of their pilgrimage, wherein they were strangers. And I have
also heard the groaning of the children of Israel, whom the Egyptians
keep in bondage; and I have remembered my covenant. Wherefore say unto
the children of Israel, I \emph{am} the \textsc{Lord}, and I will bring
you out from under the burdens of the Egyptians, and I will rid you out
of their bondage, and I will redeem you with a stretched out arm, and
with great judgments: And I will take you to me for a people, and I will
be to you a God: and ye shall know that I \emph{am} the \textsc{Lord}
your God, which bringeth you out from under the burdens of the
Egyptians. And I will bring you in unto the land, concerning the which I
did swear to give it to Abraham, to Isaac, and to Jacob; and I will give
it you for an heritage: I \emph{am} the \textsc{Lord}.

And Moses spake so unto the children of Israel: but they hearkened not
unto Moses for anguish of spirit, and for cruel bondage. And the
\textsc{Lord} spake unto Moses, saying, Go in, speak unto Pharaoh king
of Egypt, that he let the children of Israel go out of his land. And
Moses spake before the \textsc{Lord}, saying, Behold, the children of
Israel have not hearkened unto me; how then shall Pharaoh hear me, who
\emph{am} of uncircumcised lips? And the \textsc{Lord} spake unto Moses
and unto Aaron, and gave them a charge unto the children of Israel, and
unto Pharaoh king of Egypt, to bring the children of Israel out of the
land of Egypt.

These \emph{be} the heads of their fathers' houses: The sons of Reuben
the firstborn of Israel; Hanoch, and Pallu, Hezron, and Carmi: these
\emph{be} the families of Reuben. And the sons of Simeon; Jemuel, and
Jamin, and Ohad, and Jachin, and Zohar, and Shaul the son of a
Canaanitish woman: these \emph{are} the families of Simeon.

And these \emph{are} the names of the sons of Levi according to their
generations; Gershon, and Kohath, and Merari: and the years of the life
of Levi \emph{were} an hundred thirty and seven years. The sons of
Gershon; Libni, and Shimi, according to their families. And the sons of
Kohath; Amram, and Izhar, and Hebron, and Uzziel: and the years of the
life of Kohath \emph{were} an hundred thirty and three years. And the
sons of Merari; Mahali and Mushi: these \emph{are} the families of Levi
according to their generations. And Amram took him Jochebed his father's
sister to wife; and she bare him Aaron and Moses: and the years of the
life of Amram \emph{were} an hundred and thirty and seven years.

And the sons of Izhar; Korah, and Nepheg, and Zichri. And the sons of
Uzziel; Mishael, and Elzaphan, and Zithri. And Aaron took him Elisheba,
daughter of Amminadab, sister of Naashon, to wife; and she bare him
Nadab, and Abihu, Eleazar, and Ithamar. And the sons of Korah; Assir,
and Elkanah, and Abiasaph: these \emph{are} the families of the
Korhites. And Eleazar Aaron's son took him \emph{one} of the daughters
of Putiel to wife; and she bare him Phinehas: these \emph{are} the heads
of the fathers of the Levites according to their families. These
\emph{are} that Aaron and Moses, to whom the \textsc{Lord} said, Bring
out the children of Israel from the land of Egypt according to their
armies. These \emph{are} they which spake to Pharaoh king of Egypt, to
bring out the children of Israel from Egypt: these \emph{are} that Moses
and Aaron.

And it came to pass on the day \emph{when} the \textsc{Lord} spake unto
Moses in the land of Egypt, That the \textsc{Lord} spake unto Moses,
saying, I \emph{am} the \textsc{Lord}: speak thou unto Pharaoh king of
Egypt all that I say unto thee. And Moses said before the \textsc{Lord},
Behold, I \emph{am} of uncircumcised lips, and how shall Pharaoh hearken
unto me?

And the \textsc{Lord} said unto Moses, See, I have made thee a god to
Pharaoh: and Aaron thy brother shall be thy prophet. Thou shalt speak
all that I command thee: and Aaron thy brother shall speak unto Pharaoh,
that he send the children of Israel out of his land. And I will harden
Pharaoh's heart, and multiply my signs and my wonders in the land of
Egypt. But Pharaoh shall not hearken unto you, that I may lay my hand
upon Egypt, and bring forth mine armies, \emph{and} my people the
children of Israel, out of the land of Egypt by great judgments. And the
Egyptians shall know that I \emph{am} the \textsc{Lord}, when I stretch
forth mine hand upon Egypt, and bring out the children of Israel from
among them. And Moses and Aaron did as the \textsc{Lord} commanded them,
so did they. And Moses \emph{was} fourscore years old, and Aaron
fourscore and three years old, when they spake unto Pharaoh.

And the \textsc{Lord} spake unto Moses and unto Aaron, saying, When
Pharaoh shall speak unto you, saying, Shew a miracle for you: then thou
shalt say unto Aaron, Take thy rod, and cast \emph{it} before Pharaoh,
\emph{and} it shall become a serpent.

And Moses and Aaron went in unto Pharaoh, and they did so as the
\textsc{Lord} had commanded: and Aaron cast down his rod before Pharaoh,
and before his servants, and it became a serpent. Then Pharaoh also
called the wise men and the sorcerers: now the magicians of Egypt, they
also did in like manner with their enchantments. For they cast down
every man his rod, and they became serpents: but Aaron's rod swallowed
up their rods. And he hardened Pharaoh's heart, that he hearkened not
unto them; as the \textsc{Lord} had said.

And the \textsc{Lord} said unto Moses, Pharaoh's heart \emph{is}
hardened, he refuseth to let the people go. Get thee unto Pharaoh in the
morning; lo, he goeth out unto the water; and thou shalt stand by the
river's brink against he come; and the rod which was turned to a serpent
shalt thou take in thine hand. And thou shalt say unto him, The
\textsc{Lord} God of the Hebrews hath sent me unto thee, saying, Let my
people go, that they may serve me in the wilderness: and, behold,
hitherto thou wouldest not hear. Thus saith the \textsc{Lord}, In this
thou shalt know that I \emph{am} the \textsc{Lord}: behold, I will smite
with the rod that \emph{is} in mine hand upon the waters which
\emph{are} in the river, and they shall be turned to blood. And the fish
that \emph{is} in the river shall die, and the river shall stink; and
the Egyptians shall lothe to drink of the water of the river.

And the \textsc{Lord} spake unto Moses, Say unto Aaron, Take thy rod,
and stretch out thine hand upon the waters of Egypt, upon their streams,
upon their rivers, and upon their ponds, and upon all their pools of
water, that they may become blood; and \emph{that} there may be blood
throughout all the land of Egypt, both in \emph{vessels of} wood, and in
\emph{vessels of} stone. And Moses and Aaron did so, as the
\textsc{Lord} commanded; and he lifted up the rod, and smote the waters
that \emph{were} in the river, in the sight of Pharaoh, and in the sight
of his servants; and all the waters that \emph{were} in the river were
turned to blood. And the fish that \emph{was} in the river died; and the
river stank, and the Egyptians could not drink of the water of the
river; and there was blood throughout all the land of Egypt. And the
magicians of Egypt did so with their enchantments: and Pharaoh's heart
was hardened, neither did he hearken unto them; as the \textsc{Lord} had
said. And Pharaoh turned and went into his house, neither did he set his
heart to this also. And all the Egyptians digged round about the river
for water to drink; for they could not drink of the water of the river.
And seven days were fulfilled, after that the \textsc{Lord} had smitten
the river.

And the \textsc{Lord} spake unto Moses, Go unto Pharaoh, and say unto
him, Thus saith the \textsc{Lord}, Let my people go, that they may serve
me. And if thou refuse to let \emph{them} go, behold, I will smite all
thy borders with frogs: And the river shall bring forth frogs
abundantly, which shall go up and come into thine house, and into thy
bedchamber, and upon thy bed, and into the house of thy servants, and
upon thy people, and into thine ovens, and into thy kneadingtroughs: And
the frogs shall come up both on thee, and upon thy people, and upon all
thy servants.

And the \textsc{Lord} spake unto Moses, Say unto Aaron, Stretch forth
thine hand with thy rod over the streams, over the rivers, and over the
ponds, and cause frogs to come up upon the land of Egypt. And Aaron
stretched out his hand over the waters of Egypt; and the frogs came up,
and covered the land of Egypt. And the magicians did so with their
enchantments, and brought up frogs upon the land of Egypt.

Then Pharaoh called for Moses and Aaron, and said, Intreat the
\textsc{Lord}, that he may take away the frogs from me, and from my
people; and I will let the people go, that they may do sacrifice unto
the \textsc{Lord}. And Moses said unto Pharaoh, Glory over me: when
shall I intreat for thee, and for thy servants, and for thy people, to
destroy the frogs from thee and thy houses, \emph{that} they may remain
in the river only? And he said, To morrow. And he said, \emph{Be it}
according to thy word: that thou mayest know that \emph{there is} none
like unto the \textsc{Lord} our God. And the frogs shall depart from
thee, and from thy houses, and from thy servants, and from thy people;
they shall remain in the river only. And Moses and Aaron went out from
Pharaoh: and Moses cried unto the \textsc{Lord} because of the frogs
which he had brought against Pharaoh. And the \textsc{Lord} did
according to the word of Moses; and the frogs died out of the houses,
out of the villages, and out of the fields. And they gathered them
together upon heaps: and the land stank. But when Pharaoh saw that there
was respite, he hardened his heart, and hearkened not unto them; as the
\textsc{Lord} had said.

And the \textsc{Lord} said unto Moses, Say unto Aaron, Stretch out thy
rod, and smite the dust of the land, that it may become lice throughout
all the land of Egypt. And they did so; for Aaron stretched out his hand
with his rod, and smote the dust of the earth, and it became lice in
man, and in beast; all the dust of the land became lice throughout all
the land of Egypt. And the magicians did so with their enchantments to
bring forth lice, but they could not: so there were lice upon man, and
upon beast. Then the magicians said unto Pharaoh, This \emph{is} the
finger of God: and Pharaoh's heart was hardened, and he hearkened not
unto them; as the \textsc{Lord} had said.

And the \textsc{Lord} said unto Moses, Rise up early in the morning, and
stand before Pharaoh; lo, he cometh forth to the water; and say unto
him, Thus saith the \textsc{Lord}, Let my people go, that they may serve
me. Else, if thou wilt not let my people go, behold, I will send swarms
\emph{of flies} upon thee, and upon thy servants, and upon thy people,
and into thy houses: and the houses of the Egyptians shall be full of
swarms \emph{of flies,} and also the ground whereon they \emph{are. }And I
will sever in that day the land of Goshen, in which my people dwell,
that no swarms \emph{of flies} shall be there; to the end thou mayest
know that I \emph{am} the \textsc{Lord} in the midst of the earth. And
I will put a division between my people and thy people: to morrow shall
this sign be. And the \textsc{Lord} did so; and there came a grievous
swarm \emph{of flies} into the house of Pharaoh, and \emph{into} his
servants' houses, and into all the land of Egypt: the land was corrupted
by reason of the swarm \emph{of flies.}

And Pharaoh called for Moses and for Aaron, and said, Go ye, sacrifice
to your God in the land. And Moses said, It is not meet so to do; for we
shall sacrifice the abomination of the Egyptians to the \textsc{Lord}
our God: lo, shall we sacrifice the abomination of the Egyptians before
their eyes, and will they not stone us? We will go three days' journey
into the wilderness, and sacrifice to the \textsc{Lord} our God, as he
shall command us. And Pharaoh said, I will let you go, that ye may
sacrifice to the \textsc{Lord} your God in the wilderness; only ye shall
not go very far away: intreat for me. And Moses said, Behold, I go out
from thee, and I will intreat the \textsc{Lord} that the swarms \emph{of
flies} may depart from Pharaoh, from his servants, and from his people,
to morrow: but let not Pharaoh deal deceitfully any more in not letting
the people go to sacrifice to the \textsc{Lord}. And Moses went out from
Pharaoh, and intreated the \textsc{Lord}. And the \textsc{Lord} did
according to the word of Moses; and he removed the swarms \emph{of
flies} from Pharaoh, from his servants, and from his people; there
remained not one. And Pharaoh hardened his heart at this time also,
neither would he let the people go.

Then the \textsc{Lord} said unto Moses, Go in unto Pharaoh, and tell
him, Thus saith the \textsc{Lord} God of the Hebrews, Let my people go,
that they may serve me. For if thou refuse to let \emph{them} go, and
wilt hold them still, Behold, the hand of the \textsc{Lord} is upon thy
cattle which \emph{is} in the field, upon the horses, upon the asses,
upon the camels, upon the oxen, and upon the sheep: \emph{there shall
be} a very grievous murrain. And the \textsc{Lord} shall sever between
the cattle of Israel and the cattle of Egypt: and there shall nothing
die of all \emph{that is} the children's of Israel. And the
\textsc{Lord} appointed a set time, saying, To morrow the \textsc{Lord}
shall do this thing in the land. And the \textsc{Lord} did that thing on
the morrow, and all the cattle of Egypt died: but of the cattle of the
children of Israel died not one. And Pharaoh sent, and, behold, there
was not one of the cattle of the Israelites dead. And the heart of
Pharaoh was hardened, and he did not let the people go.

And the \textsc{Lord} said unto Moses and unto Aaron, Take to you
handfuls of ashes of the furnace, and let Moses sprinkle it toward the
heaven in the sight of Pharaoh. And it shall become small dust in all
the land of Egypt, and shall be a boil breaking forth \emph{with} blains
upon man, and upon beast, throughout all the land of Egypt. And they
took ashes of the furnace, and stood before Pharaoh; and Moses sprinkled
it up toward heaven; and it became a boil breaking forth \emph{with}
blains upon man, and upon beast. And the magicians could not stand
before Moses because of the boils; for the boil was upon the magicians,
and upon all the Egyptians. And the \textsc{Lord} hardened the heart of
Pharaoh, and he hearkened not unto them; as the \textsc{Lord} had spoken
unto Moses.

And the \textsc{Lord} said unto Moses, Rise up early in the morning, and
stand before Pharaoh, and say unto him, Thus saith the \textsc{Lord} God
of the Hebrews, Let my people go, that they may serve me. For I will at
this time send all my plagues upon thine heart, and upon thy servants,
and upon thy people; that thou mayest know that \emph{there is} none
like me in all the earth. For now I will stretch out my hand, that I may
smite thee and thy people with pestilence; and thou shalt be cut off
from the earth. And in very deed for this \emph{cause} have I raised
thee up, for to shew \emph{in} thee my power; and that my name may be
declared throughout all the earth. As yet exaltest thou thyself against
my people, that thou wilt not let them go? Behold, to morrow about this
time I will cause it to rain a very grievous hail, such as hath not been
in Egypt since the foundation thereof even until now. Send therefore
now, \emph{and} gather thy cattle, and all that thou hast in the field;
\emph{for upon} every man and beast which shall be found in the field,
and shall not be brought home, the hail shall come down upon them, and
they shall die. He that feared the word of the \textsc{Lord} among the
servants of Pharaoh made his servants and his cattle flee into the
houses: And he that regarded not the word of the \textsc{Lord} left his
servants and his cattle in the field.

And the \textsc{Lord} said unto Moses, Stretch forth thine hand toward
heaven, that there may be hail in all the land of Egypt, upon man, and
upon beast, and upon every herb of the field, throughout the land of
Egypt. And Moses stretched forth his rod toward heaven: and the
\textsc{Lord} sent thunder and hail, and the fire ran along upon the
ground; and the \textsc{Lord} rained hail upon the land of Egypt. So
there was hail, and fire mingled with the hail, very grievous, such as
there was none like it in all the land of Egypt since it became a
nation. And the hail smote throughout all the land of Egypt all that
\emph{was} in the field, both man and beast; and the hail smote every
herb of the field, and brake every tree of the field. Only in the land
of Goshen, where the children of Israel \emph{were,} was there no hail.

And Pharaoh sent, and called for Moses and Aaron, and said unto them, I
have sinned this time: the \textsc{Lord} \emph{is} righteous, and I and
my people \emph{are} wicked. Intreat the \textsc{Lord} (for \emph{it is}
enough) that there be no \emph{more} mighty thunderings and hail; and I
will let you go, and ye shall stay no longer. And Moses said unto him,
As soon as I am gone out of the city, I will spread abroad my hands unto
the \textsc{Lord}; \emph{and} the thunder shall cease, neither shall
there be any more hail; that thou mayest know how that the earth
\emph{is} the \textsc{Lord}'S. But as for thee and thy servants, I know
that ye will not yet fear the \textsc{Lord} God. And the flax and the
barley was smitten: for the barley was in the ear, and the flax
\emph{was} bolled. But the wheat and the rie were not smitten: for they
\emph{were} not grown up. And Moses went out of the city from Pharaoh,
and spread abroad his hands unto the \textsc{Lord}: and the thunders and
hail ceased, and the rain was not poured upon the earth. And when
Pharaoh saw that the rain and the hail and the thunders were ceased, he
sinned yet more, and hardened his heart, he and his servants. And the
heart of Pharaoh was hardened, neither would he let the children of
Israel go; as the \textsc{Lord} had spoken by Moses.

And the \textsc{Lord} said unto Moses, Go in unto Pharaoh: for I have
hardened his heart, and the heart of his servants, that I might shew
these my signs before him: And that thou mayest tell in the ears of thy
son, and of thy son's son, what things I have wrought in Egypt, and my
signs which I have done among them; that ye may know how that I
\emph{am} the \textsc{Lord}. And Moses and Aaron came in unto Pharaoh,
and said unto him, Thus saith the \textsc{Lord} God of the Hebrews, How
long wilt thou refuse to humble thyself before me? let my people go,
that they may serve me. Else, if thou refuse to let my people go,
behold, to morrow will I bring the locusts into thy coast: And they
shall cover the face of the earth, that one cannot be able to see the
earth: and they shall eat the residue of that which is escaped, which
remaineth unto you from the hail, and shall eat every tree which groweth
for you out of the field: And they shall fill thy houses, and the houses
of all thy servants, and the houses of all the Egyptians; which neither
thy fathers, nor thy fathers' fathers have seen, since the day that they
were upon the earth unto this day. And he turned himself, and went out
from Pharaoh. And Pharaoh's servants said unto him, How long shall this
man be a snare unto us? let the men go, that they may serve the
\textsc{Lord} their God: knowest thou not yet that Egypt is destroyed?
And Moses and Aaron were brought again unto Pharaoh: and he said unto
them, Go, serve the \textsc{Lord} your God: \emph{but} who \emph{are}
they that shall go? And Moses said, We will go with our young and with
our old, with our sons and with our daughters, with our flocks and with
our herds will we go; for we \emph{must hold} a feast unto the
\textsc{Lord}. And he said unto them, Let the \textsc{Lord} be so with
you, as I will let you go, and your little ones: look \emph{to it;} for
evil \emph{is} before you. Not so: go now ye \emph{that are} men, and
serve the \textsc{Lord}; for that ye did desire. And they were driven
out from Pharaoh's presence.

And the \textsc{Lord} said unto Moses, Stretch out thine hand over the
land of Egypt for the locusts, that they may come up upon the land of
Egypt, and eat every herb of the land, \emph{even} all that the hail
hath left. And Moses stretched forth his rod over the land of Egypt, and
the \textsc{Lord} brought an east wind upon the land all that day, and
all \emph{that} night; \emph{and} when it was morning, the east wind
brought the locusts. And the locusts went up over all the land of Egypt,
and rested in all the coasts of Egypt: very grievous \emph{were they;}
before them there were no such locusts as they, neither after them shall
be such. For they covered the face of the whole earth, so that the land
was darkened; and they did eat every herb of the land, and all the fruit
of the trees which the hail had left: and there remained not any green
thing in the trees, or in the herbs of the field, through all the land
of Egypt.

Then Pharaoh called for Moses and Aaron in haste; and he said, I have
sinned against the \textsc{Lord} your God, and against you. Now
therefore forgive, I pray thee, my sin only this once, and intreat the
\textsc{Lord} your God, that he may take away from me this death only.
And he went out from Pharaoh, and intreated the \textsc{Lord}. And the
\textsc{Lord} turned a mighty strong west wind, which took away the
locusts, and cast them into the Red sea; there remained not one locust
in all the coasts of Egypt. But the \textsc{Lord} hardened Pharaoh's
heart, so that he would not let the children of Israel go.

And the \textsc{Lord} said unto Moses, Stretch out thine hand toward
heaven, that there may be darkness over the land of Egypt, even darkness
\emph{which} may be felt. And Moses stretched forth his hand toward
heaven; and there was a thick darkness in all the land of Egypt three
days: They saw not one another, neither rose any from his place for
three days: but all the children of Israel had light in their dwellings.

And Pharaoh called unto Moses, and said, Go ye, serve the \textsc{Lord};
only let your flocks and your herds be stayed: let your little ones also
go with you. And Moses said, Thou must give us also sacrifices and burnt
offerings, that we may sacrifice unto the \textsc{Lord} our God. Our
cattle also shall go with us; there shall not an hoof be left behind;
for thereof must we take to serve the \textsc{Lord} our God; and we know
not with what we must serve the \textsc{Lord}, until we come thither.

But the \textsc{Lord} hardened Pharaoh's heart, and he would not let
them go. And Pharaoh said unto him, Get thee from me, take heed to
thyself, see my face no more; for in \emph{that} day thou seest my face
thou shalt die. And Moses said, Thou hast spoken well, I will see thy
face again no more.

And the \textsc{Lord} said unto Moses, Yet will I bring one plague
\emph{more} upon Pharaoh, and upon Egypt; afterwards he will let you go
hence: when he shall let \emph{you} go, he shall surely thrust you out
hence altogether. Speak now in the ears of the people, and let every man
borrow of his neighbour, and every woman of her neighbour, jewels of
silver, and jewels of gold. And the \textsc{Lord} gave the people favour
in the sight of the Egyptians. Moreover the man Moses \emph{was} very
great in the land of Egypt, in the sight of Pharaoh's servants, and in
the sight of the people. And Moses said, Thus saith the \textsc{Lord},
About midnight will I go out into the midst of Egypt: And all the
firstborn in the land of Egypt shall die, from the firstborn of Pharaoh
that sitteth upon his throne, even unto the firstborn of the maidservant
that \emph{is} behind the mill; and all the firstborn of beasts. And
there shall be a great cry throughout all the land of Egypt, such as
there was none like it, nor shall be like it any more. But against any
of the children of Israel shall not a dog move his tongue, against man
or beast: that ye may know how that the \textsc{Lord} doth put a
difference between the Egyptians and Israel. And all these thy servants
shall come down unto me, and bow down themselves unto me, saying, Get
thee out, and all the people that follow thee: and after that I will go
out. And he went out from Pharaoh in a great anger. And the
\textsc{Lord} said unto Moses, Pharaoh shall not hearken unto you; that
my wonders may be multiplied in the land of Egypt. And Moses and Aaron
did all these wonders before Pharaoh: and the \textsc{Lord} hardened
Pharaoh's heart, so that he would not let the children of Israel go out
of his land.

And the \textsc{Lord} spake unto Moses and Aaron in the land of Egypt,
saying, This month \emph{shall be} unto you the beginning of months: it
\emph{shall be} the first month of the year to you.

Speak ye unto all the congregation of Israel, saying, In the tenth
\emph{day} of this month they shall take to them every man a lamb,
according to the house of \emph{their} fathers, a lamb for an house: And
if the household be too little for the lamb, let him and his neighbour
next unto his house take \emph{it} according to the number of the souls;
every man according to his eating shall make your count for the lamb.
Your lamb shall be without blemish, a male of the first year: ye shall
take \emph{it} out from the sheep, or from the goats: And ye shall keep
it up until the fourteenth day of the same month: and the whole assembly
of the congregation of Israel shall kill it in the evening. And they
shall take of the blood, and strike \emph{it} on the two side posts and
on the upper door post of the houses, wherein they shall eat it. And
they shall eat the flesh in that night, roast with fire, and unleavened
bread; \emph{and} with bitter \emph{herbs} they shall eat it. Eat not of
it raw, nor sodden at all with water, but roast \emph{with} fire; his
head with his legs, and with the purtenance thereof. And ye shall let
nothing of it remain until the morning; and that which remaineth of it
until the morning ye shall burn with fire.

And thus shall ye eat it; \emph{with} your loins girded, your shoes on
your feet, and your staff in your hand; and ye shall eat it in haste: it
\emph{is} the \textsc{Lord}'S passover. For I will pass through the land
of Egypt this night, and will smite all the firstborn in the land of
Egypt, both man and beast; and against all the gods of Egypt I will
execute judgment: I \emph{am} the \textsc{Lord}. And the blood shall be
to you for a token upon the houses where ye \emph{are:} and when I see
the blood, I will pass over you, and the plague shall not be upon you to
destroy \emph{you,} when I smite the land of Egypt. And this day shall
be unto you for a memorial; and ye shall keep it a feast to the
\textsc{Lord} throughout your generations; ye shall keep it a feast by
an ordinance for ever. Seven days shall ye eat unleavened bread; even
the first day ye shall put away leaven out of your houses: for whosoever
eateth leavened bread from the first day until the seventh day, that
soul shall be cut off from Israel. And in the first day \emph{there
shall be} an holy convocation, and in the seventh day there shall be an
holy convocation to you; no manner of work shall be done in them, save
\emph{that} which every man must eat, that only may be done of you. And
ye shall observe \emph{the feast of} unleavened bread; for in this
selfsame day have I brought your armies out of the land of Egypt:
therefore shall ye observe this day in your generations by an ordinance
for ever.

In the first \emph{month,} on the fourteenth day of the month at even,
ye shall eat unleavened bread, until the one and twentieth day of the
month at even. Seven days shall there be no leaven found in your houses:
for whosoever eateth that which is leavened, even that soul shall be cut
off from the congregation of Israel, whether he be a stranger, or born
in the land. Ye shall eat nothing leavened; in all your habitations
shall ye eat unleavened bread.

Then Moses called for all the elders of Israel, and said unto them, Draw
out and take you a lamb according to your families, and kill the
passover. And ye shall take a bunch of hyssop, and dip \emph{it} in the
blood that \emph{is} in the bason, and strike the lintel and the two
side posts with the blood that \emph{is} in the bason; and none of you
shall go out at the door of his house until the morning. For the
\textsc{Lord} will pass through to smite the Egyptians; and when he
seeth the blood upon the lintel, and on the two side posts, the
\textsc{Lord} will pass over the door, and will not suffer the destroyer
to come in unto your houses to smite \emph{you. }And ye shall observe
this thing for an ordinance to thee and to thy sons for ever. And it
shall come to pass, when ye be come to the land which the \textsc{Lord}
will give you, according as he hath promised, that ye shall keep this
service. And it shall come to pass, when your children shall say unto
you, What mean ye by this service? That ye shall say, It \emph{is} the
sacrifice of the \textsc{Lord}'S passover, who passed over the houses of
the children of Israel in Egypt, when he smote the Egyptians, and
delivered our houses. And the people bowed the head and worshipped. And
the children of Israel went away, and did as the \textsc{Lord} had
commanded Moses and Aaron, so did they.

And it came to pass, that at midnight the \textsc{Lord} smote all the
firstborn in the land of Egypt, from the firstborn of Pharaoh that sat
on his throne unto the firstborn of the captive that \emph{was} in the
dungeon; and all the firstborn of cattle. And Pharaoh rose up in the
night, he, and all his servants, and all the Egyptians; and there was a
great cry in Egypt; for \emph{there was} not a house where \emph{there
was} not one dead.

And he called for Moses and Aaron by night, and said, Rise up,
\emph{and} get you forth from among my people, both ye and the children
of Israel; and go, serve the \textsc{Lord}, as ye have said. Also take
your flocks and your herds, as ye have said, and be gone; and bless me
also. And the Egyptians were urgent upon the people, that they might
send them out of the land in haste; for they said, We \emph{be} all dead
\emph{men. }And the people took their dough before it was leavened,
their kneadingtroughs being bound up in their clothes upon their
shoulders. And the children of Israel did according to the word of
Moses; and they borrowed of the Egyptians jewels of silver, and jewels
of gold, and raiment: And the \textsc{Lord} gave the people favour in
the sight of the Egyptians, so that they lent unto them \emph{such things as
they required.} And they spoiled the Egyptians.

And the children of Israel journeyed from Rameses to Succoth, about six
hundred thousand on foot \emph{that were} men, beside children. And a
mixed multitude went up also with them; and flocks, and herds,
\emph{even} very much cattle. And they baked unleavened cakes of the
dough which they brought forth out of Egypt, for it was not leavened;
because they were thrust out of Egypt, and could not tarry, neither had
they prepared for themselves any victual.

Now the sojourning of the children of Israel, who dwelt in Egypt,
\emph{was} four hundred and thirty years. And it came to pass at the end
of the four hundred and thirty years, even the selfsame day it came to
pass, that all the hosts of the \textsc{Lord} went out from the land of
Egypt. It \emph{is} a night to be much observed unto the \textsc{Lord}
for bringing them out from the land of Egypt: this \emph{is} that night
of the \textsc{Lord} to be observed of all the children of Israel in
their generations.

And the \textsc{Lord} said unto Moses and Aaron, This \emph{is} the
ordinance of the passover: There shall no stranger eat thereof: But
every man's servant that is bought for money, when thou hast circumcised
him, then shall he eat thereof. A foreigner and an hired servant shall
not eat thereof. In one house shall it be eaten; thou shalt not carry
forth ought of the flesh abroad out of the house; neither shall ye break
a bone thereof. All the congregation of Israel shall keep it. And when a
stranger shall sojourn with thee, and will keep the passover to the
\textsc{Lord}, let all his males be circumcised, and then let him come
near and keep it; and he shall be as one that is born in the land: for
no uncircumcised person shall eat thereof. One law shall be to him that
is homeborn, and unto the stranger that sojourneth among you. Thus did
all the children of Israel; as the \textsc{Lord} commanded Moses and
Aaron, so did they. And it came to pass the selfsame day, \emph{that}
the \textsc{Lord} did bring the children of Israel out of the land of
Egypt by their armies.

And the \textsc{Lord} spake unto Moses, saying, Sanctify unto me all the
firstborn, whatsoever openeth the womb among the children of Israel,
\emph{both} of man and of beast: it \emph{is} mine.

And Moses said unto the people, Remember this day, in which ye came out
from Egypt, out of the house of bondage; for by strength of hand the
\textsc{Lord} brought you out from this \emph{place:} there shall no
leavened bread be eaten. This day came ye out in the month Abib.

And it shall be when the \textsc{Lord} shall bring thee into the land of
the Canaanites, and the Hittites, and the Amorites, and the Hivites, and
the Jebusites, which he sware unto thy fathers to give thee, a land
flowing with milk and honey, that thou shalt keep this service in this
month. Seven days thou shalt eat unleavened bread, and in the seventh
day \emph{shall be} a feast to the \textsc{Lord}. Unleavened bread shall
be eaten seven days; and there shall no leavened bread be seen with
thee, neither shall there be leaven seen with thee in all thy quarters.

And thou shalt shew thy son in that day, saying, \emph{This is done} because
of that \emph{which} the \textsc{Lord} did unto me when I came forth out
of Egypt. And it shall be for a sign unto thee upon thine hand, and for
a memorial between thine eyes, that the \textsc{Lord}'S law may be in
thy mouth: for with a strong hand hath the \textsc{Lord} brought thee
out of Egypt. Thou shalt therefore keep this ordinance in his season
from year to year.

And it shall be when the \textsc{Lord} shall bring thee into the land of
the Canaanites, as he sware unto thee and to thy fathers, and shall give
it thee, That thou shalt set apart unto the \textsc{Lord} all that
openeth the matrix, and every firstling that cometh of a beast which
thou hast; the males \emph{shall be} the \textsc{Lord}'S. And every
firstling of an ass thou shalt redeem with a lamb; and if thou wilt not
redeem it, then thou shalt break his neck: and all the firstborn of man
among thy children shalt thou redeem.

And it shall be when thy son asketh thee in time to come, saying, What
\emph{is} this? that thou shalt say unto him, By strength of hand the
\textsc{Lord} brought us out from Egypt, from the house of bondage: And
it came to pass, when Pharaoh would hardly let us go, that the
\textsc{Lord} slew all the firstborn in the land of Egypt, both the
firstborn of man, and the firstborn of beast: therefore I sacrifice to
the \textsc{Lord} all that openeth the matrix, being males; but all the
firstborn of my children I redeem. And it shall be for a token upon
thine hand, and for frontlets between thine eyes: for by strength of
hand the \textsc{Lord} brought us forth out of Egypt.

And it came to pass, when Pharaoh had let the people go, that God led
them not \emph{through} the way of the land of the Philistines, although
that \emph{was} near; for God said, Lest peradventure the people repent
when they see war, and they return to Egypt: But God led the people
about, \emph{through} the way of the wilderness of the Red sea: and the
children of Israel went up harnessed out of the land of Egypt. And Moses
took the bones of Joseph with him: for he had straitly sworn the
children of Israel, saying, God will surely visit you; and ye shall
carry up my bones away hence with you.

And they took their journey from Succoth, and encamped in Etham, in the
edge of the wilderness. And the \textsc{Lord} went before them by day in
a pillar of a cloud, to lead them the way; and by night in a pillar of
fire, to give them light; to go by day and night: He took not away the
pillar of the cloud by day, nor the pillar of fire by night, \emph{from}
before the people.

And the \textsc{Lord} spake unto Moses, saying, Speak unto the children
of Israel, that they turn and encamp before Pi-hahiroth, between Migdol
and the sea, over against Baal-zephon: before it shall ye encamp by the
sea. For Pharaoh will say of the children of Israel, They \emph{are}
entangled in the land, the wilderness hath shut them in. And I will
harden Pharaoh's heart, that he shall follow after them; and I will be
honoured upon Pharaoh, and upon all his host; that the Egyptians may
know that I \emph{am} the \textsc{Lord}. And they did so.

And it was told the king of Egypt that the people fled: and the heart of
Pharaoh and of his servants was turned against the people, and they
said, Why have we done this, that we have let Israel go from serving us?
And he made ready his chariot, and took his people with him: And he took
six hundred chosen chariots, and all the chariots of Egypt, and captains
over every one of them. And the \textsc{Lord} hardened the heart of
Pharaoh king of Egypt, and he pursued after the children of Israel: and
the children of Israel went out with an high hand. But the Egyptians
pursued after them, all the horses \emph{and} chariots of Pharaoh, and
his horsemen, and his army, and overtook them encamping by the sea,
beside Pi-hahiroth, before Baal-zephon.

And when Pharaoh drew nigh, the children of Israel lifted up their eyes,
and, behold, the Egyptians marched after them; and they were sore
afraid: and the children of Israel cried out unto the \textsc{Lord}. And
they said unto Moses, Because \emph{there were} no graves in Egypt, hast
thou taken us away to die in the wilderness? wherefore hast thou dealt
thus with us, to carry us forth out of Egypt? \emph{Is} not this the
word that we did tell thee in Egypt, saying, Let us alone, that we may
serve the Egyptians? For \emph{it had been} better for us to serve the
Egyptians, than that we should die in the wilderness.

And Moses said unto the people, Fear ye not, stand still, and see the
salvation of the \textsc{Lord}, which he will shew to you to day: for
the Egyptians whom ye have seen to day, ye shall see them again no more
for ever. The \textsc{Lord} shall fight for you, and ye shall hold your
peace.

And the \textsc{Lord} said unto Moses, Wherefore criest thou unto me?
speak unto the children of Israel, that they go forward: But lift thou
up thy rod, and stretch out thine hand over the sea, and divide it: and
the children of Israel shall go on dry \emph{ground} through the midst
of the sea. And I, behold, I will harden the hearts of the Egyptians,
and they shall follow them: and I will get me honour upon Pharaoh, and
upon all his host, upon his chariots, and upon his horsemen. And the
Egyptians shall know that I \emph{am} the \textsc{Lord}, when I have
gotten me honour upon Pharaoh, upon his chariots, and upon his horsemen.

And the angel of God, which went before the camp of Israel, removed and
went behind them; and the pillar of the cloud went from before their
face, and stood behind them: And it came between the camp of the
Egyptians and the camp of Israel; and it was a cloud and darkness
\emph{to them,} but it gave light by night \emph{to these:} so that the
one came not near the other all the night. And Moses stretched out his
hand over the sea; and the \textsc{Lord} caused the sea to go
\emph{back} by a strong east wind all that night, and made the sea dry
\emph{land,} and the waters were divided. And the children of Israel
went into the midst of the sea upon the dry \emph{ground:} and the
waters \emph{were} a wall unto them on their right hand, and on their
left.

And the Egyptians pursued, and went in after them to the midst of the
sea, \emph{even} all Pharaoh's horses, his chariots, and his horsemen.
And it came to pass, that in the morning watch the \textsc{Lord} looked
unto the host of the Egyptians through the pillar of fire and of the
cloud, and troubled the host of the Egyptians, And took off their
chariot wheels, that they drave them heavily: so that the Egyptians
said, Let us flee from the face of Israel; for the \textsc{Lord}
fighteth for them against the Egyptians.

And the \textsc{Lord} said unto Moses, Stretch out thine hand over the
sea, that the waters may come again upon the Egyptians, upon their
chariots, and upon their horsemen. And Moses stretched forth his hand
over the sea, and the sea returned to his strength when the morning
appeared; and the Egyptians fled against it; and the \textsc{Lord}
overthrew the Egyptians in the midst of the sea. And the waters
returned, and covered the chariots, and the horsemen, \emph{and} all the
host of Pharaoh that came into the sea after them; there remained not so
much as one of them. But the children of Israel walked upon dry
\emph{land} in the midst of the sea; and the waters \emph{were} a wall
unto them on their right hand, and on their left. Thus the \textsc{Lord}
saved Israel that day out of the hand of the Egyptians; and Israel saw
the Egyptians dead upon the sea shore. And Israel saw that great work
which the \textsc{Lord} did upon the Egyptians: and the people feared
the \textsc{Lord}, and believed the \textsc{Lord}, and his servant
Moses.

Then sang Moses and the children of Israel this song unto the
\textsc{Lord}, and spake, saying, I will sing unto the \textsc{Lord},
for he hath triumphed gloriously: the horse and his rider hath he thrown
into the sea. The \textsc{Lord} \emph{is} my strength and song, and he
is become my salvation: he \emph{is} my God, and I will prepare him an
habitation; my father's God, and I will exalt him. The \textsc{Lord}
\emph{is} a man of war: the \textsc{Lord} \emph{is} his name. Pharaoh's
chariots and his host hath he cast into the sea: his chosen captains
also are drowned in the Red sea. The depths have covered them: they sank
into the bottom as a stone. Thy right hand, O \textsc{Lord}, is become
glorious in power: thy right hand, O \textsc{Lord}, hath dashed in
pieces the enemy. And in the greatness of thine excellency thou hast
overthrown them that rose up against thee: thou sentest forth thy wrath,
\emph{which} consumed them as stubble. And with the blast of thy
nostrils the waters were gathered together, the floods stood upright as
an heap, \emph{and} the depths were congealed in the heart of the sea.
The enemy said, I will pursue, I will overtake, I will divide the spoil;
my lust shall be satisfied upon them; I will draw my sword, my hand
shall destroy them. Thou didst blow with thy wind, the sea covered them:
they sank as lead in the mighty waters. Who \emph{is} like unto thee, O
\textsc{Lord}, among the gods? who \emph{is} like thee, glorious in
holiness, fearful \emph{in} praises, doing wonders? Thou stretchedst out
thy right hand, the earth swallowed them. Thou in thy mercy hast led
forth the people \emph{which} thou hast redeemed: thou hast guided
\emph{them} in thy strength unto thy holy habitation. The people shall
hear, \emph{and} be afraid: sorrow shall take hold on the inhabitants of
Palestina. Then the dukes of Edom shall be amazed; the mighty men of
Moab, trembling shall take hold upon them; all the inhabitants of Canaan
shall melt away. Fear and dread shall fall upon them; by the greatness
of thine arm they shall be \emph{as} still as a stone; till thy people
pass over, O \textsc{Lord}, till the people pass over, \emph{which} thou
hast purchased. Thou shalt bring them in, and plant them in the mountain
of thine inheritance, \emph{in} the place, O \textsc{Lord}, \emph{which}
thou hast made for thee to dwell in, in the Sanctuary, O Lord,
\emph{which} thy hands have established. The \textsc{Lord} shall reign
for ever and ever. For the horse of Pharaoh went in with his chariots
and with his horsemen into the sea, and the \textsc{Lord} brought again
the waters of the sea upon them; but the children of Israel went on dry
\emph{land} in the midst of the sea.

And Miriam the prophetess, the sister of Aaron, took a timbrel in her
hand; and all the women went out after her with timbrels and with
dances. And Miriam answered them, Sing ye to the \textsc{Lord}, for he
hath triumphed gloriously; the horse and his rider hath he thrown into
the sea. So Moses brought Israel from the Red sea, and they went out
into the wilderness of Shur; and they went three days in the wilderness,
and found no water.

And when they came to Marah, they could not drink of the waters of
Marah, for they \emph{were} bitter: therefore the name of it was called
Marah. And the people murmured against Moses, saying, What shall we
drink? And he cried unto the \textsc{Lord}; and the \textsc{Lord} shewed
him a tree, \emph{which} when he had cast into the waters, the waters
were made sweet: there he made for them a statute and an ordinance, and
there he proved them, And said, If thou wilt diligently hearken to the
voice of the \textsc{Lord} thy God, and wilt do that which is right in
his sight, and wilt give ear to his commandments, and keep all his
statutes, I will put none of these diseases upon thee, which I have
brought upon the Egyptians: for I \emph{am} the \textsc{Lord} that
healeth thee.

And they came to Elim, where \emph{were} twelve wells of water, and
threescore and ten palm trees: and they encamped there by the waters.

And they took their journey from Elim, and all the congregation of the
children of Israel came unto the wilderness of Sin, which \emph{is}
between Elim and Sinai, on the fifteenth day of the second month after
their departing out of the land of Egypt. And the whole congregation of
the children of Israel murmured against Moses and Aaron in the
wilderness: And the children of Israel said unto them, Would to God we
had died by the hand of the \textsc{Lord} in the land of Egypt, when we
sat by the flesh pots, \emph{and} when we did eat bread to the full; for
ye have brought us forth into this wilderness, to kill this whole
assembly with hunger.

Then said the \textsc{Lord} unto Moses, Behold, I will rain bread from
heaven for you; and the people shall go out and gather a certain rate
every day, that I may prove them, whether they will walk in my law, or
no. And it shall come to pass, that on the sixth day they shall prepare
\emph{that} which they bring in; and it shall be twice as much as they
gather daily. And Moses and Aaron said unto all the children of Israel,
At even, then ye shall know that the \textsc{Lord} hath brought you out
from the land of Egypt: And in the morning, then ye shall see the glory
of the \textsc{Lord}; for that he heareth your murmurings against the
\textsc{Lord}: and what \emph{are} we, that ye murmur against us? And
Moses said, \emph{This shall be,} when the \textsc{Lord} shall give you
in the evening flesh to eat, and in the morning bread to the full; for
that the \textsc{Lord} heareth your murmurings which ye murmur against
him: and what \emph{are} we? your murmurings \emph{are} not against us,
but against the \textsc{Lord}.

And Moses spake unto Aaron, Say unto all the congregation of the
children of Israel, Come near before the \textsc{Lord}: for he hath
heard your murmurings. And it came to pass, as Aaron spake unto the
whole congregation of the children of Israel, that they looked toward
the wilderness, and, behold, the glory of the \textsc{Lord} appeared in
the cloud.

And the \textsc{Lord} spake unto Moses, saying, I have heard the
murmurings of the children of Israel: speak unto them, saying, At even
ye shall eat flesh, and in the morning ye shall be filled with bread;
and ye shall know that I \emph{am} the \textsc{Lord} your God. And it
came to pass, that at even the quails came up, and covered the camp: and
in the morning the dew lay round about the host. And when the dew that
lay was gone up, behold, upon the face of the wilderness \emph{there
lay} a small round thing, \emph{as} small as the hoar frost on the
ground. And when the children of Israel saw \emph{it,} they said one to
another, It \emph{is} manna: for they wist not what it \emph{was.} And
Moses said unto them, This \emph{is} the bread which the \textsc{Lord}
hath given you to eat.

This \emph{is} the thing which the \textsc{Lord} hath commanded, Gather
of it every man \emph{according to} his eating, an omer for every man,
according to the number of your persons; take ye every man for
\emph{them} which \emph{are} in his tents. And the children of Israel
did so, and gathered, some more, some less. And when they did mete
\emph{it} with an omer, he that gathered much had nothing over, and he
that gathered little had no lack; they gathered every man according to
his eating. And Moses said, Let no man leave of it till the morning.
Notwithstanding they hearkened not unto Moses; but some of them left of
it until the morning, and it bred worms, and stank: and Moses was wroth
with them. And they gathered it every morning, every man according to
his eating: and when the sun waxed hot, it melted.

And it came to pass, \emph{that} on the sixth day they gathered twice as
much bread, two omers for one \emph{man:} and all the rulers of the
congregation came and told Moses. And he said unto them, This \emph{is
that} which the \textsc{Lord} hath said, To morrow \emph{is} the rest of
the holy sabbath unto the \textsc{Lord}: bake \emph{that} which ye will
bake \emph{to day,} and seethe that ye will seethe; and that which
remaineth over lay up for you to be kept until the morning. And they
laid it up till the morning, as Moses bade: and it did not stink,
neither was there any worm therein. And Moses said, Eat that to day; for
to day \emph{is} a sabbath unto the \textsc{Lord}: to day ye shall not
find it in the field. Six days ye shall gather it; but on the seventh
day, \emph{which is} the sabbath, in it there shall be none.

And it came to pass, \emph{that} there went out \emph{some} of the
people on the seventh day for to gather, and they found none. And the
\textsc{Lord} said unto Moses, How long refuse ye to keep my
commandments and my laws? See, for that the \textsc{Lord} hath given you
the sabbath, therefore he giveth you on the sixth day the bread of two
days; abide ye every man in his place, let no man go out of his place on
the seventh day. So the people rested on the seventh day. And the house
of Israel called the name thereof Manna: and it \emph{was} like
coriander seed, white; and the taste of it \emph{was} like wafers
\emph{made} with honey.

And Moses said, This \emph{is} the thing which the \textsc{Lord}
commandeth, Fill an omer of it to be kept for your generations; that
they may see the bread wherewith I have fed you in the wilderness, when
I brought you forth from the land of Egypt. And Moses said unto Aaron,
Take a pot, and put an omer full of manna therein, and lay it up before
the \textsc{Lord}, to be kept for your generations. As the \textsc{Lord}
commanded Moses, so Aaron laid it up before the Testimony, to be kept.
And the children of Israel did eat manna forty years, until they came to
a land inhabited; they did eat manna, until they came unto the borders
of the land of Canaan. Now an omer \emph{is} the tenth \emph{part} of an
ephah.

And all the congregation of the children of Israel journeyed from the
wilderness of Sin, after their journeys, according to the commandment of
the \textsc{Lord}, and pitched in Rephidim: and \emph{there was} no
water for the people to drink. Wherefore the people did chide with
Moses, and said, Give us water that we may drink. And Moses said unto
them, Why chide ye with me? wherefore do ye tempt the \textsc{Lord}? And
the people thirsted there for water; and the people murmured against
Moses, and said, Wherefore \emph{is} this \emph{that} thou hast brought
us up out of Egypt, to kill us and our children and our cattle with
thirst? And Moses cried unto the \textsc{Lord}, saying, What shall I do
unto this people? they be almost ready to stone me. And the
\textsc{Lord} said unto Moses, Go on before the people, and take with
thee of the elders of Israel; and thy rod, wherewith thou smotest the
river, take in thine hand, and go. Behold, I will stand before thee
there upon the rock in Horeb; and thou shalt smite the rock, and there
shall come water out of it, that the people may drink. And Moses did so
in the sight of the elders of Israel. And he called the name of the
place Massah, and Meribah, because of the chiding of the children of
Israel, and because they tempted the \textsc{Lord}, saying, Is the
\textsc{Lord} among us, or not?

Then came Amalek, and fought with Israel in Rephidim. And Moses said
unto Joshua, Choose us out men, and go out, fight with Amalek: to morrow
I will stand on the top of the hill with the rod of God in mine hand. So
Joshua did as Moses had said to him, and fought with Amalek: and Moses,
Aaron, and Hur went up to the top of the hill. And it came to pass, when
Moses held up his hand, that Israel prevailed: and when he let down his
hand, Amalek prevailed. But Moses' hands \emph{were} heavy; and they
took a stone, and put \emph{it} under him, and he sat thereon; and Aaron
and Hur stayed up his hands, the one on the one side, and the other on
the other side; and his hands were steady until the going down of the
sun. And Joshua discomfited Amalek and his people with the edge of the
sword. And the \textsc{Lord} said unto Moses, Write this \emph{for} a
memorial in a book, and rehearse \emph{it} in the ears of Joshua: for I
will utterly put out the remembrance of Amalek from under heaven. And
Moses built an altar, and called the name of it Jehovah-nissi: For he
said, Because the \textsc{Lord} hath sworn \emph{that} the \textsc{Lord}
\emph{will have} war with Amalek from generation to generation.

When Jethro, the priest of Midian, Moses' father in law, heard of all
that God had done for Moses, and for Israel his people, \emph{and} that
the \textsc{Lord} had brought Israel out of Egypt; Then Jethro, Moses'
father in law, took Zipporah, Moses' wife, after he had sent her back,
And her two sons; of which the name of the one \emph{was} Gershom; for
he said, I have been an alien in a strange land: And the name of the
other \emph{was} Eliezer; for the God of my father, \emph{said he, was}
mine help, and delivered me from the sword of Pharaoh: And Jethro,
Moses' father in law, came with his sons and his wife unto Moses into
the wilderness, where he encamped at the mount of God: And he said unto
Moses, I thy father in law Jethro am come unto thee, and thy wife, and
her two sons with her.

And Moses went out to meet his father in law, and did obeisance, and
kissed him; and they asked each other of \emph{their} welfare; and they
came into the tent. And Moses told his father in law all that the
\textsc{Lord} had done unto Pharaoh and to the Egyptians for Israel's
sake, \emph{and} all the travail that had come upon them by the way, and
\emph{how} the \textsc{Lord} delivered them. And Jethro rejoiced for all
the goodness which the \textsc{Lord} had done to Israel, whom he had
delivered out of the hand of the Egyptians. And Jethro said, Blessed
\emph{be} the \textsc{Lord}, who hath delivered you out of the hand of
the Egyptians, and out of the hand of Pharaoh, who hath delivered the
people from under the hand of the Egyptians. Now I know that the
\textsc{Lord} \emph{is} greater than all gods: for in the thing wherein
they dealt proudly \emph{he was} above them. And Jethro, Moses' father
in law, took a burnt offering and sacrifices for God: and Aaron came,
and all the elders of Israel, to eat bread with Moses' father in law
before God.

And it came to pass on the morrow, that Moses sat to judge the people:
and the people stood by Moses from the morning unto the evening. And
when Moses' father in law saw all that he did to the people, he said,
What \emph{is} this thing that thou doest to the people? why sittest
thou thyself alone, and all the people stand by thee from morning unto
even? And Moses said unto his father in law, Because the people come
unto me to enquire of God: When they have a matter, they come unto me;
and I judge between one and another, and I do make \emph{them} know the
statutes of God, and his laws. And Moses' father in law said unto him,
The thing that thou doest \emph{is} not good. Thou wilt surely wear
away, both thou, and this people that \emph{is} with thee: for this
thing \emph{is} too heavy for thee; thou art not able to perform it
thyself alone. Hearken now unto my voice, I will give thee counsel, and
God shall be with thee: Be thou for the people to God-ward, that thou
mayest bring the causes unto God: And thou shalt teach them ordinances
and laws, and shalt shew them the way wherein they must walk, and the
work that they must do. Moreover thou shalt provide out of all the
people able men, such as fear God, men of truth, hating covetousness;
and place \emph{such} over them, \emph{to be} rulers of thousands,
\emph{and} rulers of hundreds, rulers of fifties, and rulers of tens:
And let them judge the people at all seasons: and it shall be,
\emph{that} every great matter they shall bring unto thee, but every
small matter they shall judge: so shall it be easier for thyself, and
they shall bear \emph{the burden} with thee. If thou shalt do this thing, and
God command thee \emph{so,} then thou shalt be able to endure, and all
this people shall also go to their place in peace. So Moses hearkened to
the voice of his father in law, and did all that he had said. And Moses
chose able men out of all Israel, and made them heads over the people,
rulers of thousands, rulers of hundreds, rulers of fifties, and rulers
of tens. And they judged the people at all seasons: the hard causes they
brought unto Moses, but every small matter they judged themselves.

And Moses let his father in law depart; and he went his way into his own
land.

In the third month, when the children of Israel were gone forth out of
the land of Egypt, the same day came they \emph{into} the wilderness of
Sinai. For they were departed from Rephidim, and were come \emph{to} the
desert of Sinai, and had pitched in the wilderness; and there Israel
camped before the mount. And Moses went up unto God, and the
\textsc{Lord} called unto him out of the mountain, saying, Thus shalt
thou say to the house of Jacob, and tell the children of Israel; Ye have
seen what I did unto the Egyptians, and \emph{how} I bare you on eagles'
wings, and brought you unto myself. Now therefore, if ye will obey my
voice indeed, and keep my covenant, then ye shall be a peculiar treasure
unto me above all people: for all the earth \emph{is} mine: And ye shall
be unto me a kingdom of priests, and an holy nation. These \emph{are}
the words which thou shalt speak unto the children of Israel.

And Moses came and called for the elders of the people, and laid before
their faces all these words which the \textsc{Lord} commanded him. And
all the people answered together, and said, All that the \textsc{Lord}
hath spoken we will do. And Moses returned the words of the people unto
the \textsc{Lord}. And the \textsc{Lord} said unto Moses, Lo, I come
unto thee in a thick cloud, that the people may hear when I speak with
thee, and believe thee for ever. And Moses told the words of the people
unto the \textsc{Lord}.

And the \textsc{Lord} said unto Moses, Go unto the people, and sanctify
them to day and to morrow, and let them wash their clothes, And be ready
against the third day: for the third day the \textsc{Lord} will come
down in the sight of all the people upon mount Sinai. And thou shalt set
bounds unto the people round about, saying, Take heed to yourselves,
\emph{that ye} go \emph{not} up into the mount, or touch the border of
it: whosoever toucheth the mount shall be surely put to death: There
shall not an hand touch it, but he shall surely be stoned, or shot
through; whether \emph{it be} beast or man, it shall not live: when the
trumpet soundeth long, they shall come up to the mount.

And Moses went down from the mount unto the people, and sanctified the
people; and they washed their clothes. And he said unto the people, Be
ready against the third day: come not at \emph{your} wives.

And it came to pass on the third day in the morning, that there were
thunders and lightnings, and a thick cloud upon the mount, and the voice
of the trumpet exceeding loud; so that all the people that \emph{was} in
the camp trembled. And Moses brought forth the people out of the camp to
meet with God; and they stood at the nether part of the mount. And mount
Sinai was altogether on a smoke, because the \textsc{Lord} descended
upon it in fire: and the smoke thereof ascended as the smoke of a
furnace, and the whole mount quaked greatly. And when the voice of the
trumpet sounded long, and waxed louder and louder, Moses spake, and God
answered him by a voice. And the \textsc{Lord} came down upon mount
Sinai, on the top of the mount: and the \textsc{Lord} called Moses
\emph{up} to the top of the mount; and Moses went up. And the
\textsc{Lord} said unto Moses, Go down, charge the people, lest they
break through unto the \textsc{Lord} to gaze, and many of them perish.
And let the priests also, which come near to the \textsc{Lord}, sanctify
themselves, lest the \textsc{Lord} break forth upon them. And Moses said
unto the \textsc{Lord}, The people cannot come up to mount Sinai: for
thou chargedst us, saying, Set bounds about the mount, and sanctify it.
And the \textsc{Lord} said unto him, Away, get thee down, and thou shalt
come up, thou, and Aaron with thee: but let not the priests and the
people break through to come up unto the \textsc{Lord}, lest he break
forth upon them. So Moses went down unto the people, and spake unto
them.

And God spake all these words, saying, I \emph{am} the \textsc{Lord} thy
God, which have brought thee out of the land of Egypt, out of the house
of bondage. Thou shalt have no other gods before me. Thou shalt not make
unto thee any graven image, or any likeness \emph{of any thing} that
\emph{is} in heaven above, or that \emph{is} in the earth beneath, or
that \emph{is} in the water under the earth: Thou shalt not bow down
thyself to them, nor serve them: for I the \textsc{Lord} thy God
\emph{am} a jealous God, visiting the iniquity of the fathers upon the
children unto the third and fourth \emph{generation} of them that hate
me; And shewing mercy unto thousands of them that love me, and keep my
commandments. Thou shalt not take the name of the \textsc{Lord} thy God
in vain; for the \textsc{Lord} will not hold him guiltless that taketh
his name in vain. Remember the sabbath day, to keep it holy. Six days
shalt thou labour, and do all thy work: But the seventh day \emph{is}
the sabbath of the \textsc{Lord} thy God: \emph{in it} thou shalt not do
any work, thou, nor thy son, nor thy daughter, thy manservant, nor thy
maidservant, nor thy cattle, nor thy stranger that \emph{is} within thy
gates: For \emph{in} six days the \textsc{Lord} made heaven and earth,
the sea, and all that in them \emph{is,} and rested the seventh day:
wherefore the \textsc{Lord} blessed the sabbath day, and hallowed it.

Honour thy father and thy mother: that thy days may be long upon the
land which the \textsc{Lord} thy God giveth thee. Thou shalt not kill.
Thou shalt not commit adultery. Thou shalt not steal. Thou shalt not
bear false witness against thy neighbour. Thou shalt not covet thy
neighbour's house, thou shalt not covet thy neighbour's wife, nor his
manservant, nor his maidservant, nor his ox, nor his ass, nor any thing
that \emph{is} thy neighbour's.

And all the people saw the thunderings, and the lightnings, and the
noise of the trumpet, and the mountain smoking: and when the people saw
\emph{it,} they removed, and stood afar off. And they said unto Moses,
Speak thou with us, and we will hear: but let not God speak with us,
lest we die. And Moses said unto the people, Fear not: for God is come
to prove you, and that his fear may be before your faces, that ye sin
not. And the people stood afar off, and Moses drew near unto the thick
darkness where God \emph{was.}

And the \textsc{Lord} said unto Moses, Thus thou shalt say unto the
children of Israel, Ye have seen that I have talked with you from
heaven. Ye shall not make with me gods of silver, neither shall ye make
unto you gods of gold.

An altar of earth thou shalt make unto me, and shalt sacrifice thereon
thy burnt offerings, and thy peace offerings, thy sheep, and thine oxen:
in all places where I record my name I will come unto thee, and I will
bless thee. And if thou wilt make me an altar of stone, thou shalt not
build it of hewn stone: for if thou lift up thy tool upon it, thou hast
polluted it. Neither shalt thou go up by steps unto mine altar, that thy
nakedness be not discovered thereon.

Now these \emph{are} the judgments which thou shalt set before them. If
thou buy an Hebrew servant, six years he shall serve: and in the seventh
he shall go out free for nothing. If he came in by himself, he shall go
out by himself: if he were married, then his wife shall go out with him.
If his master have given him a wife, and she have born him sons or
daughters; the wife and her children shall be her master's, and he shall
go out by himself. And if the servant shall plainly say, I love my
master, my wife, and my children; I will not go out free: Then his
master shall bring him unto the judges; he shall also bring him to the
door, or unto the door post; and his master shall bore his ear through
with an aul; and he shall serve him for ever.

And if a man sell his daughter to be a maidservant, she shall not go out
as the menservants do. If she please not her master, who hath betrothed
her to himself, then shall he let her be redeemed: to sell her unto a
strange nation he shall have no power, seeing he hath dealt deceitfully
with her. And if he have betrothed her unto his son, he shall deal with
her after the manner of daughters. If he take him another \emph{wife;}
her food, her raiment, and her duty of marriage, shall he not diminish.
And if he do not these three unto her, then shall she go out free
without money.

He that smiteth a man, so that he die, shall be surely put to death. And
if a man lie not in wait, but God deliver \emph{him} into his hand; then
I will appoint thee a place whither he shall flee. But if a man come
presumptuously upon his neighbour, to slay him with guile; thou shalt
take him from mine altar, that he may die.

And he that smiteth his father, or his mother, shall be surely put to
death.

And he that stealeth a man, and selleth him, or if he be found in his
hand, he shall surely be put to death.

And he that curseth his father, or his mother, shall surely be put to
death.

And if men strive together, and one smite another with a stone, or with
\emph{his} fist, and he die not, but keepeth \emph{his} bed: If he rise
again, and walk abroad upon his staff, then shall he that smote
\emph{him} be quit: only he shall pay \emph{for} the loss of his time,
and shall cause \emph{him} to be thoroughly healed.

And if a man smite his servant, or his maid, with a rod, and he die
under his hand; he shall be surely punished. Notwithstanding, if he
continue a day or two, he shall not be punished: for he \emph{is} his
money.

If men strive, and hurt a woman with child, so that her fruit depart
\emph{from her,} and yet no mischief follow: he shall be surely
punished, according as the woman's husband will lay upon him; and he
shall pay as the judges \emph{determine. }And if \emph{any} mischief
follow, then thou shalt give life for life, Eye for eye, tooth for
tooth, hand for hand, foot for foot, Burning for burning, wound for
wound, stripe for stripe.

And if a man smite the eye of his servant, or the eye of his maid, that
it perish; he shall let him go free for his eye's sake. And if he smite
out his manservant's tooth, or his maidservant's tooth; he shall let him
go free for his tooth's sake.

If an ox gore a man or a woman, that they die: then the ox shall be
surely stoned, and his flesh shall not be eaten; but the owner of the ox
\emph{shall be} quit. But if the ox were wont to push with his horn in
time past, and it hath been testified to his owner, and he hath not kept
him in, but that he hath killed a man or a woman; the ox shall be
stoned, and his owner also shall be put to death. If there be laid on
him a sum of money, then he shall give for the ransom of his life
whatsoever is laid upon him. Whether he have gored a son, or have gored
a daughter, according to this judgment shall it be done unto him. If the
ox shall push a manservant or a maidservant; he shall give unto their
master thirty shekels of silver, and the ox shall be stoned.

And if a man shall open a pit, or if a man shall dig a pit, and not
cover it, and an ox or an ass fall therein; The owner of the pit shall
make \emph{it} good, \emph{and} give money unto the owner of them; and
the dead \emph{beast} shall be his.

And if one man's ox hurt another's, that he die; then they shall sell
the live ox, and divide the money of it; and the dead \emph{ox} also
they shall divide. Or if it be known that the ox hath used to push in
time past, and his owner hath not kept him in; he shall surely pay ox
for ox; and the dead shall be his own.

If a man shall steal an ox, or a sheep, and kill it, or sell it; he
shall restore five oxen for an ox, and four sheep for a sheep.

If a thief be found breaking up, and be smitten that he die, \emph{there
shall} no blood \emph{be shed} for him. If the sun be risen upon him,
\emph{there shall be} blood \emph{shed} for him; \emph{for} he should
make full restitution; if he have nothing, then he shall be sold for his
theft. If the theft be certainly found in his hand alive, whether it be
ox, or ass, or sheep; he shall restore double.

If a man shall cause a field or vineyard to be eaten, and shall put in
his beast, and shall feed in another man's field; of the best of his own
field, and of the best of his own vineyard, shall he make restitution.

If fire break out, and catch in thorns, so that the stacks of corn, or
the standing corn, or the field, be consumed \emph{therewith;} he that
kindled the fire shall surely make restitution.

If a man shall deliver unto his neighbour money or stuff to keep, and it
be stolen out of the man's house; if the thief be found, let him pay
double. If the thief be not found, then the master of the house shall be
brought unto the judges, \emph{to see} whether he have put his hand unto
his neighbour's goods. For all manner of trespass, \emph{whether it be}
for ox, for ass, for sheep, for raiment, \emph{or} for any manner of
lost thing, which \emph{another} challengeth to be his, the cause of
both parties shall come before the judges; \emph{and} whom the judges
shall condemn, he shall pay double unto his neighbour. If a man deliver
unto his neighbour an ass, or an ox, or a sheep, or any beast, to keep;
and it die, or be hurt, or driven away, no man seeing \emph{it:
}\emph{Then} shall an oath of the \textsc{Lord} be between them both,
that he hath not put his hand unto his neighbour's goods; and the owner
of it shall accept \emph{thereof,} and he shall not make \emph{it} good.
And if it be stolen from him, he shall make restitution unto the owner
thereof. If it be torn in pieces, \emph{then} let him bring it
\emph{for} witness, \emph{and} he shall not make good that which was
torn.

And if a man borrow \emph{ought} of his neighbour, and it be hurt, or
die, the owner thereof \emph{being} not with it, he shall surely make
\emph{it} good. \emph{But} if the owner thereof \emph{be} with it, he
shall not make \emph{it} good: if it \emph{be} an hired \emph{thing,} it
came for his hire.

And if a man entice a maid that is not betrothed, and lie with her, he
shall surely endow her to be his wife. If her father utterly refuse to
give her unto him, he shall pay money according to the dowry of virgins.

Thou shalt not suffer a witch to live.

Whosoever lieth with a beast shall surely be put to death.

He that sacrificeth unto \emph{any} god, save unto the \textsc{Lord}
only, he shall be utterly destroyed.

Thou shalt neither vex a stranger, nor oppress him: for ye were
strangers in the land of Egypt.

Ye shall not afflict any widow, or fatherless child. If thou afflict
them in any wise, and they cry at all unto me, I will surely hear their
cry; And my wrath shall wax hot, and I will kill you with the sword; and
your wives shall be widows, and your children fatherless.

If thou lend money to \emph{any of} my people \emph{that is} poor by
thee, thou shalt not be to him as an usurer, neither shalt thou lay upon
him usury. If thou at all take thy neighbour's raiment to pledge, thou
shalt deliver it unto him by that the sun goeth down: For that \emph{is}
his covering only, it \emph{is} his raiment for his skin: wherein shall
he sleep? and it shall come to pass, when he crieth unto me, that I will
hear; for I \emph{am} gracious.

Thou shalt not revile the gods, nor curse the ruler of thy people.

Thou shalt not delay \emph{to offer} the first of thy ripe fruits, and
of thy liquors: the firstborn of thy sons shalt thou give unto me.
Likewise shalt thou do with thine oxen, \emph{and} with thy sheep: seven
days it shall be with his dam; on the eighth day thou shalt give it me.

And ye shall be holy men unto me: neither shall ye eat \emph{any} flesh
\emph{that is} torn of beasts in the field; ye shall cast it to the
dogs.

Thou shalt not raise a false report: put not thine hand with the wicked
to be an unrighteous witness.

Thou shalt not follow a multitude to \emph{do} evil; neither shalt thou
speak in a cause to decline after many to wrest \emph{judgment:}

Neither shalt thou countenance a poor man in his cause.

If thou meet thine enemy's ox or his ass going astray, thou shalt surely
bring it back to him again. If thou see the ass of him that hateth thee
lying under his burden, and wouldest forbear to help him, thou shalt
surely help with him. Thou shalt not wrest the judgment of thy poor in
his cause. Keep thee far from a false matter; and the innocent and
righteous slay thou not: for I will not justify the wicked.

And thou shalt take no gift: for the gift blindeth the wise, and
perverteth the words of the righteous.

Also thou shalt not oppress a stranger: for ye know the heart of a
stranger, seeing ye were strangers in the land of Egypt. And six years
thou shalt sow thy land, and shalt gather in the fruits thereof: But the
seventh \emph{year} thou shalt let it rest and lie still; that the poor
of thy people may eat: and what they leave the beasts of the field shall
eat. In like manner thou shalt deal with thy vineyard, \emph{and} with
thy oliveyard. Six days thou shalt do thy work, and on the seventh day
thou shalt rest: that thine ox and thine ass may rest, and the son of
thy handmaid, and the stranger, may be refreshed. And in all
\emph{things} that I have said unto you be circumspect: and make no
mention of the name of other gods, neither let it be heard out of thy
mouth.

Three times thou shalt keep a feast unto me in the year. Thou shalt keep
the feast of unleavened bread: (thou shalt eat unleavened bread seven
days, as I commanded thee, in the time appointed of the month Abib; for
in it thou camest out from Egypt: and none shall appear before me
empty:) And the feast of harvest, the firstfruits of thy labours, which
thou hast sown in the field: and the feast of ingathering, \emph{which
is} in the end of the year, when thou hast gathered in thy labours out
of the field. Three times in the year all thy males shall appear before
the Lord GOD. Thou shalt not offer the blood of my sacrifice with
leavened bread; neither shall the fat of my sacrifice remain until the
morning. The first of the firstfruits of thy land thou shalt bring into
the house of the \textsc{Lord} thy God. Thou shalt not seethe a kid in
his mother's milk.

Behold, I send an Angel before thee, to keep thee in the way, and to
bring thee into the place which I have prepared. Beware of him, and obey
his voice, provoke him not; for he will not pardon your transgressions:
for my name \emph{is} in him. But if thou shalt indeed obey his voice,
and do all that I speak; then I will be an enemy unto thine enemies, and
an adversary unto thine adversaries. For mine Angel shall go before
thee, and bring thee in unto the Amorites, and the Hittites, and the
Perizzites, and the Canaanites, and the Hivites, and the Jebusites: and
I will cut them off. Thou shalt not bow down to their gods, nor serve
them, nor do after their works: but thou shalt utterly overthrow them,
and quite break down their images. And ye shall serve the \textsc{Lord}
your God, and he shall bless thy bread, and thy water; and I will take
sickness away from the midst of thee.

There shall nothing cast their young, nor be barren, in thy land: the
number of thy days I will fulfil. I will send my fear before thee, and
will destroy all the people to whom thou shalt come, and I will make all
thine enemies turn their backs unto thee. And I will send hornets before
thee, which shall drive out the Hivite, the Canaanite, and the Hittite,
from before thee. I will not drive them out from before thee in one
year; lest the land become desolate, and the beast of the field multiply
against thee. By little and little I will drive them out from before
thee, until thou be increased, and inherit the land. And I will set thy
bounds from the Red sea even unto the sea of the Philistines, and from
the desert unto the river: for I will deliver the inhabitants of the
land into your hand; and thou shalt drive them out before thee. Thou
shalt make no covenant with them, nor with their gods. They shall not
dwell in thy land, lest they make thee sin against me: for if thou serve
their gods, it will surely be a snare unto thee.

And he said unto Moses, Come up unto the \textsc{Lord}, thou, and Aaron,
Nadab, and Abihu, and seventy of the elders of Israel; and worship ye
afar off. And Moses alone shall come near the \textsc{Lord}: but they
shall not come nigh; neither shall the people go up with him.

And Moses came and told the people all the words of the \textsc{Lord},
and all the judgments: and all the people answered with one voice, and
said, All the words which the \textsc{Lord} hath said will we do. And
Moses wrote all the words of the \textsc{Lord}, and rose up early in the
morning, and builded an altar under the hill, and twelve pillars,
according to the twelve tribes of Israel. And he sent young men of the
children of Israel, which offered burnt offerings, and sacrificed peace
offerings of oxen unto the \textsc{Lord}. And Moses took half of the
blood, and put \emph{it} in basons; and half of the blood he sprinkled
on the altar. And he took the book of the covenant, and read in the
audience of the people: and they said, All that the \textsc{Lord} hath
said will we do, and be obedient. And Moses took the blood, and
sprinkled \emph{it} on the people, and said, Behold the blood of the
covenant, which the \textsc{Lord} hath made with you concerning all
these words.

Then went up Moses, and Aaron, Nadab, and Abihu, and seventy of the
elders of Israel: And they saw the God of Israel: and \emph{there was}
under his feet as it were a paved work of a sapphire stone, and as it
were the body of heaven in \emph{his} clearness. And upon the nobles of
the children of Israel he laid not his hand: also they saw God, and did
eat and drink.

And the \textsc{Lord} said unto Moses, Come up to me into the mount, and
be there: and I will give thee tables of stone, and a law, and
commandments which I have written; that thou mayest teach them. And
Moses rose up, and his minister Joshua: and Moses went up into the mount
of God. And he said unto the elders, Tarry ye here for us, until we come
again unto you: and, behold, Aaron and Hur \emph{are} with you: if any
man have any matters to do, let him come unto them. And Moses went up
into the mount, and a cloud covered the mount. And the glory of the
\textsc{Lord} abode upon mount Sinai, and the cloud covered it six days:
and the seventh day he called unto Moses out of the midst of the cloud.
And the sight of the glory of the \textsc{Lord} \emph{was} like
devouring fire on the top of the mount in the eyes of the children of
Israel. And Moses went into the midst of the cloud, and gat him up into
the mount: and Moses was in the mount forty days and forty nights.

And the \textsc{Lord} spake unto Moses, saying, Speak unto the children
of Israel, that they bring me an offering: of every man that giveth it
willingly with his heart ye shall take my offering. And this \emph{is}
the offering which ye shall take of them; gold, and silver, and brass,
And blue, and purple, and scarlet, and fine linen, and goats \emph{hair,
}And rams' skins dyed red, and badgers' skins, and shittim wood, Oil for
the light, spices for anointing oil, and for sweet incense, Onyx stones,
and stones to be set in the ephod, and in the breastplate. And let them
make me a sanctuary; that I may dwell among them. According to all that
I shew thee, \emph{after} the pattern of the tabernacle, and the pattern
of all the instruments thereof, even so shall ye make \emph{it.}

And they shall make an ark \emph{of} shittim wood: two cubits and a half
\emph{shall be} the length thereof, and a cubit and a half the breadth
thereof, and a cubit and a half the height thereof. And thou shalt
overlay it with pure gold, within and without shalt thou overlay it, and
shalt make upon it a crown of gold round about. And thou shalt cast four
rings of gold for it, and put \emph{them} in the four corners thereof;
and two rings \emph{shall be} in the one side of it, and two rings in
the other side of it. And thou shalt make staves \emph{of} shittim wood,
and overlay them with gold. And thou shalt put the staves into the rings
by the sides of the ark, that the ark may be borne with them. The staves
shall be in the rings of the ark: they shall not be taken from it. And
thou shalt put into the ark the testimony which I shall give thee. And
thou shalt make a mercy seat \emph{of} pure gold: two cubits and a half
\emph{shall be} the length thereof, and a cubit and a half the breadth
thereof. And thou shalt make two cherubims \emph{of} gold, \emph{of}
beaten work shalt thou make them, in the two ends of the mercy seat. And
make one cherub on the one end, and the other cherub on the other end:
\emph{even} of the mercy seat shall ye make the cherubims on the two
ends thereof. And the cherubim shall stretch forth \emph{their} wings on
high, covering the mercy seat with their wings, and their faces \emph{shall
look} one to another; toward the mercy seat shall the faces of the
cherubims be. And thou shalt put the mercy seat above upon the ark; and
in the ark thou shalt put the testimony that I shall give thee. And
there I will meet with thee, and I will commune with thee from above the
mercy seat, from between the two cherubims which \emph{are} upon the ark
of the testimony, of all \emph{things} which I will give thee in
commandment unto the children of Israel.

Thou shalt also make a table \emph{of} shittim wood: two cubits
\emph{shall be} the length thereof, and a cubit the breadth thereof, and
a cubit and a half the height thereof. And thou shalt overlay it with
pure gold, and make thereto a crown of gold round about. And thou shalt
make unto it a border of an hand breadth round about, and thou shalt
make a golden crown to the border thereof round about. And thou shalt
make for it four rings of gold, and put the rings in the four corners
that \emph{are} on the four feet thereof. Over against the border shall
the rings be for places of the staves to bear the table. And thou shalt
make the staves \emph{of} shittim wood, and overlay them with gold, that
the table may be borne with them. And thou shalt make the dishes
thereof, and spoons thereof, and covers thereof, and bowls thereof, to
cover withal: \emph{of} pure gold shalt thou make them. And thou shalt
set upon the table shewbread before me alway.

And thou shalt make a candlestick \emph{of} pure gold: \emph{of} beaten
work shall the candlestick be made: his shaft, and his branches, his
bowls, his knops, and his flowers, shall be of the same. And six
branches shall come out of the sides of it; three branches of the
candlestick out of the one side, and three branches of the candlestick
out of the other side: Three bowls made like unto almonds, \emph{with} a
knop and a flower in one branch; and three bowls made like almonds in
the other branch, \emph{with} a knop and a flower: so in the six
branches that come out of the candlestick. And in the candlestick
\emph{shall be} four bowls made like unto almonds, \emph{with} their
knops and their flowers. And \emph{there shall be} a knop under two
branches of the same, and a knop under two branches of the same, and a
knop under two branches of the same, according to the six branches that
proceed out of the candlestick. Their knops and their branches shall be
of the same: all it \emph{shall be} one beaten work \emph{of} pure gold.
And thou shalt make the seven lamps thereof: and they shall light the
lamps thereof, that they may give light over against it. And the tongs
thereof, and the snuffdishes thereof, \emph{shall be of} pure gold. \emph{Of}
a talent of pure gold shall he make it, with all these vessels. And look
that thou make \emph{them} after their pattern, which was shewed thee in
the mount.

Moreover thou shalt make the tabernacle \emph{with} ten curtains
\emph{of} fine twined linen, and blue, and purple, and scarlet:
\emph{with} cherubims of cunning work shalt thou make them. The length
of one curtain \emph{shall be} eight and twenty cubits, and the breadth
of one curtain four cubits: and every one of the curtains shall have one
measure. The five curtains shall be coupled together one to another; and
\emph{other} five curtains \emph{shall be} coupled one to another. And
thou shalt make loops of blue upon the edge of the one curtain from the
selvedge in the coupling; and likewise shalt thou make in the uttermost
edge of \emph{another} curtain, in the coupling of the second. Fifty
loops shalt thou make in the one curtain, and fifty loops shalt thou
make in the edge of the curtain that is in the coupling of the second;
that the loops may take hold one of another. And thou shalt make fifty
taches of gold, and couple the curtains together with the taches: and it
shall be one tabernacle.

And thou shalt make curtains \emph{of} goats' \emph{hair} to be a
covering upon the tabernacle: eleven curtains shalt thou make. The
length of one curtain \emph{shall be} thirty cubits, and the breadth of
one curtain four cubits: and the eleven curtains \emph{shall be all} of
one measure. And thou shalt couple five curtains by themselves, and six
curtains by themselves, and shalt double the sixth curtain in the
forefront of the tabernacle. And thou shalt make fifty loops on the edge
of the one curtain \emph{that is} outmost in the coupling, and fifty
loops in the edge of the curtain which coupleth the second. And thou
shalt make fifty taches of brass, and put the taches into the loops, and
couple the tent together, that it may be one. And the remnant that
remaineth of the curtains of the tent, the half curtain that remaineth,
shall hang over the backside of the tabernacle. And a cubit on the one
side, and a cubit on the other side of that which remaineth in the
length of the curtains of the tent, it shall hang over the sides of the
tabernacle on this side and on that side, to cover it. And thou shalt
make a covering for the tent \emph{of} rams' skins dyed red, and a
covering above \emph{of} badgers skins.

And thou shalt make boards for the tabernacle \emph{of} shittim wood
standing up. Ten cubits \emph{shall be} the length of a board, and a
cubit and a half \emph{shall be} the breadth of one board. Two tenons
\emph{shall there be} in one board, set in order one against another:
thus shalt thou make for all the boards of the tabernacle. And thou
shalt make the boards for the tabernacle, twenty boards on the south
side southward. And thou shalt make forty sockets of silver under the
twenty boards; two sockets under one board for his two tenons, and two
sockets under another board for his two tenons. And for the second side
of the tabernacle on the north side \emph{there shall be} twenty boards: And
their forty sockets \emph{of} silver; two sockets under one board, and
two sockets under another board. And for the sides of the tabernacle
westward thou shalt make six boards. And two boards shalt thou make for
the corners of the tabernacle in the two sides. And they shall be
coupled together beneath, and they shall be coupled together above the
head of it unto one ring: thus shall it be for them both; they shall be
for the two corners. And they shall be eight boards, and their sockets
\emph{of} silver, sixteen sockets; two sockets under one board, and two
sockets under another board.

And thou shalt make bars \emph{of} shittim wood; five for the boards of
the one side of the tabernacle, And five bars for the boards of the
other side of the tabernacle, and five bars for the boards of the side
of the tabernacle, for the two sides westward. And the middle bar in the
midst of the boards shall reach from end to end. And thou shalt overlay
the boards with gold, and make their rings \emph{of} gold \emph{for}
places for the bars: and thou shalt overlay the bars with gold. And thou
shalt rear up the tabernacle according to the fashion thereof which was
shewed thee in the mount.

And thou shalt make a vail \emph{of} blue, and purple, and scarlet, and
fine twined linen of cunning work: with cherubims shall it be made: And
thou shalt hang it upon four pillars of shittim \emph{wood} overlaid
with gold: their hooks \emph{shall be of} gold, upon the four sockets of
silver.

And thou shalt hang up the vail under the taches, that thou mayest bring
in thither within the vail the ark of the testimony: and the vail shall
divide unto you between the holy \emph{place} and the most holy. And
thou shalt put the mercy seat upon the ark of the testimony in the most
holy \emph{place. }And thou shalt set the table without the vail, and
the candlestick over against the table on the side of the tabernacle
toward the south: and thou shalt put the table on the north side. And
thou shalt make an hanging for the door of the tent, \emph{of} blue, and
purple, and scarlet, and fine twined linen, wrought with needlework. And
thou shalt make for the hanging five pillars \emph{of} shittim
\emph{wood,} and overlay them with gold, \emph{and} their hooks
\emph{shall be of} gold: and thou shalt cast five sockets of brass for
them.

And thou shalt make an altar \emph{of} shittim wood, five cubits long,
and five cubits broad; the altar shall be foursquare: and the height
thereof \emph{shall be} three cubits. And thou shalt make the horns of
it upon the four corners thereof: his horns shall be of the same: and
thou shalt overlay it with brass. And thou shalt make his pans to
receive his ashes, and his shovels, and his basons, and his fleshhooks,
and his firepans: all the vessels thereof thou shalt make \emph{of}
brass. And thou shalt make for it a grate of network \emph{of} brass;
and upon the net shalt thou make four brasen rings in the four corners
thereof. And thou shalt put it under the compass of the altar beneath,
that the net may be even to the midst of the altar. And thou shalt make
staves for the altar, staves \emph{of} shittim wood, and overlay them
with brass. And the staves shall be put into the rings, and the staves
shall be upon the two sides of the altar, to bear it. Hollow with boards
shalt thou make it: as it was shewed thee in the mount, so shall they
make \emph{it.}

And thou shalt make the court of the tabernacle: for the south side
southward \emph{there shall be} hangings for the court \emph{of} fine
twined linen of an hundred cubits long for one side: And the twenty
pillars thereof and their twenty sockets \emph{shall be of} brass; the
hooks of the pillars and their fillets \emph{shall be of} silver. And
likewise for the north side in length \emph{there shall be} hangings of
an hundred \emph{cubits} long, and his twenty pillars and their twenty
sockets \emph{of} brass; the hooks of the pillars and their fillets
\emph{of} silver.

And \emph{for} the breadth of the court on the west side \emph{shall be}
hangings of fifty cubits: their pillars ten, and their sockets ten. And
the breadth of the court on the east side eastward \emph{shall be} fifty
cubits. The hangings of one side \emph{of the gate shall be} fifteen cubits:
their pillars three, and their sockets three. And on the other side
\emph{shall be} hangings fifteen \emph{cubits:} their pillars three, and
their sockets three.

And for the gate of the court \emph{shall be} an hanging of twenty
cubits, \emph{of} blue, and purple, and scarlet, and fine twined linen,
wrought with needlework: \emph{and} their pillars \emph{shall be} four,
and their sockets four. All the pillars round about the court
\emph{shall be} filleted with silver; their hooks \emph{shall be of}
silver, and their sockets of brass.

The length of the court \emph{shall be} an hundred cubits, and the
breadth fifty every where, and the height five cubits \emph{of} fine
twined linen, and their sockets \emph{of} brass. All the vessels of the
tabernacle in all the service thereof, and all the pins thereof, and all
the pins of the court, \emph{shall be of} brass.

And thou shalt command the children of Israel, that they bring thee pure
oil olive beaten for the light, to cause the lamp to burn always. In the
tabernacle of the congregation without the vail, which \emph{is} before
the testimony, Aaron and his sons shall order it from evening to morning
before the \textsc{Lord}: \emph{it shall be} a statute for ever unto
their generations on the behalf of the children of Israel.

And take thou unto thee Aaron thy brother, and his sons with him, from
among the children of Israel, that he may minister unto me in the
priest's office, \emph{even} Aaron, Nadab and Abihu, Eleazar and
Ithamar, Aaron's sons. And thou shalt make holy garments for Aaron thy
brother for glory and for beauty. And thou shalt speak unto all
\emph{that are} wise hearted, whom I have filled with the spirit of
wisdom, that they may make Aaron's garments to consecrate him, that he
may minister unto me in the priest's office. And these \emph{are} the
garments which they shall make; a breastplate, and an ephod, and a robe,
and a broidered coat, a mitre, and a girdle: and they shall make holy
garments for Aaron thy brother, and his sons, that he may minister unto
me in the priest's office. And they shall take gold, and blue, and
purple, and scarlet, and fine linen.

And they shall make the ephod \emph{of} gold, \emph{of} blue, and
\emph{of} purple, \emph{of} scarlet, and fine twined linen, with cunning
work. It shall have the two shoulderpieces thereof joined at the two
edges thereof; and \emph{so} it shall be joined together. And the
curious girdle of the ephod, which \emph{is} upon it, shall be of the
same, according to the work thereof; \emph{even of} gold, \emph{of}
blue, and purple, and scarlet, and fine twined linen. And thou shalt
take two onyx stones, and grave on them the names of the children of
Israel: Six of their names on one stone, and \emph{the other} six names
of the rest on the other stone, according to their birth. With the work
of an engraver in stone, \emph{like} the engravings of a signet, shalt
thou engrave the two stones with the names of the children of Israel:
thou shalt make them to be set in ouches of gold. And thou shalt put the
two stones upon the shoulders of the ephod \emph{for} stones of memorial
unto the children of Israel: and Aaron shall bear their names before the
\textsc{Lord} upon his two shoulders for a memorial.

And thou shalt make ouches \emph{of} gold; And two chains \emph{of} pure
gold at the ends; \emph{of} wreathen work shalt thou make them, and
fasten the wreathen chains to the ouches.

And thou shalt make the breastplate of judgment with cunning work; after
the work of the ephod thou shalt make it; \emph{of} gold, \emph{of}
blue, and \emph{of} purple, and \emph{of} scarlet, and \emph{of} fine
twined linen, shalt thou make it. Foursquare it shall be \emph{being}
doubled; a span \emph{shall be} the length thereof, and a span
\emph{shall be} the breadth thereof. And thou shalt set in it settings
of stones, \emph{even} four rows of stones: \emph{the first} row
\emph{shall be} a sardius, a topaz, and a carbuncle: \emph{this shall
be} the first row. And the second row \emph{shall be} an emerald, a
sapphire, and a diamond. And the third row a ligure, an agate, and an
amethyst. And the fourth row a beryl, and an onyx, and a jasper: they
shall be set in gold in their inclosings. And the stones shall be with
the names of the children of Israel, twelve, according to their names,
\emph{like} the engravings of a signet; every one with his name shall
they be according to the twelve tribes.

And thou shalt make upon the breastplate chains at the ends \emph{of}
wreathen work \emph{of} pure gold. And thou shalt make upon the
breastplate two rings of gold, and shalt put the two rings on the two
ends of the breastplate. And thou shalt put the two wreathen
\emph{chains} of gold in the two rings \emph{which are} on the ends of
the breastplate. And \emph{the other} two ends of the two wreathen
\emph{chains} thou shalt fasten in the two ouches, and put \emph{them}
on the shoulderpieces of the ephod before it.

And thou shalt make two rings of gold, and thou shalt put them upon the
two ends of the breastplate in the border thereof, which \emph{is} in
the side of the ephod inward. And two \emph{other} rings of gold thou
shalt make, and shalt put them on the two sides of the ephod underneath,
toward the forepart thereof, over against the \emph{other} coupling
thereof, above the curious girdle of the ephod. And they shall bind the
breastplate by the rings thereof unto the rings of the ephod with a lace
of blue, that \emph{it} may be above the curious girdle of the ephod,
and that the breastplate be not loosed from the ephod. And Aaron shall
bear the names of the children of Israel in the breastplate of judgment
upon his heart, when he goeth in unto the holy \emph{place,} for a
memorial before the \textsc{Lord} continually.

And thou shalt put in the breastplate of judgment the Urim and the
Thummim; and they shall be upon Aaron's heart, when he goeth in before
the \textsc{Lord}: and Aaron shall bear the judgment of the children of
Israel upon his heart before the \textsc{Lord} continually.

And thou shalt make the robe of the ephod all \emph{of} blue. And there
shall be an hole in the top of it, in the midst thereof: it shall have a
binding of woven work round about the hole of it, as it were the hole of
an habergeon, that it be not rent.

And \emph{beneath} upon the hem of it thou shalt make pomegranates
\emph{of} blue, and \emph{of} purple, and \emph{of} scarlet, round about
the hem thereof; and bells of gold between them round about: A golden
bell and a pomegranate, a golden bell and a pomegranate, upon the hem of
the robe round about. And it shall be upon Aaron to minister: and his
sound shall be heard when he goeth in unto the holy \emph{place} before
the \textsc{Lord}, and when he cometh out, that he die not.

And thou shalt make a plate \emph{of} pure gold, and grave upon it,
\emph{like} the engravings of a signet, HOLINESS TO THE \textsc{Lord}.
And thou shalt put it on a blue lace, that it may be upon the mitre;
upon the forefront of the mitre it shall be. And it shall be upon
Aaron's forehead, that Aaron may bear the iniquity of the holy things,
which the children of Israel shall hallow in all their holy gifts; and
it shall be always upon his forehead, that they may be accepted before
the \textsc{Lord}.

And thou shalt embroider the coat of fine linen, and thou shalt make the
mitre \emph{of} fine linen, and thou shalt make the girdle \emph{of}
needlework.

And for Aaron's sons thou shalt make coats, and thou shalt make for them
girdles, and bonnets shalt thou make for them, for glory and for beauty.
And thou shalt put them upon Aaron thy brother, and his sons with him;
and shalt anoint them, and consecrate them, and sanctify them, that they
may minister unto me in the priest's office. And thou shalt make them
linen breeches to cover their nakedness; from the loins even unto the
thighs they shall reach: And they shall be upon Aaron, and upon his
sons, when they come in unto the tabernacle of the congregation, or when
they come near unto the altar to minister in the holy \emph{place;} that
they bear not iniquity, and die: \emph{it shall be} a statute for ever
unto him and his seed after him.

And this \emph{is} the thing that thou shalt do unto them to hallow
them, to minister unto me in the priest's office: Take one young
bullock, and two rams without blemish, And unleavened bread, and cakes
unleavened tempered with oil, and wafers unleavened anointed with oil:
\emph{of} wheaten flour shalt thou make them. And thou shalt put them
into one basket, and bring them in the basket, with the bullock and the
two rams. And Aaron and his sons thou shalt bring unto the door of the
tabernacle of the congregation, and shalt wash them with water. And thou
shalt take the garments, and put upon Aaron the coat, and the robe of
the ephod, and the ephod, and the breastplate, and gird him with the
curious girdle of the ephod: And thou shalt put the mitre upon his head,
and put the holy crown upon the mitre. Then shalt thou take the
anointing oil, and pour \emph{it} upon his head, and anoint him. And
thou shalt bring his sons, and put coats upon them. And thou shalt gird
them with girdles, Aaron and his sons, and put the bonnets on them: and
the priest's office shall be theirs for a perpetual statute: and thou
shalt consecrate Aaron and his sons. And thou shalt cause a bullock to
be brought before the tabernacle of the congregation: and Aaron and his
sons shall put their hands upon the head of the bullock. And thou shalt
kill the bullock before the \textsc{Lord}, \emph{by} the door of the
tabernacle of the congregation. And thou shalt take of the blood of the
bullock, and put \emph{it} upon the horns of the altar with thy finger,
and pour all the blood beside the bottom of the altar. And thou shalt
take all the fat that covereth the inwards, and the caul \emph{that is}
above the liver, and the two kidneys, and the fat that is upon them, and
burn \emph{them} upon the altar. But the flesh of the bullock, and his
skin, and his dung, shalt thou burn with fire without the camp: it
\emph{is} a sin offering.

Thou shalt also take one ram; and Aaron and his sons shall put their
hands upon the head of the ram. And thou shalt slay the ram, and thou
shalt take his blood, and sprinkle \emph{it} round about upon the altar.
And thou shalt cut the ram in pieces, and wash the inwards of him, and
his legs, and put \emph{them} unto his pieces, and unto his head. And
thou shalt burn the whole ram upon the altar: it \emph{is} a burnt
offering unto the \textsc{Lord}: it \emph{is} a sweet savour, an
offering made by fire unto the \textsc{Lord}.

And thou shalt take the other ram; and Aaron and his sons shall put
their hands upon the head of the ram. Then shalt thou kill the ram, and
take of his blood, and put \emph{it} upon the tip of the right ear of
Aaron, and upon the tip of the right ear of his sons, and upon the thumb
of their right hand, and upon the great toe of their right foot, and
sprinkle the blood upon the altar round about. And thou shalt take of
the blood that \emph{is} upon the altar, and of the anointing oil, and
sprinkle \emph{it} upon Aaron, and upon his garments, and upon his sons,
and upon the garments of his sons with him: and he shall be hallowed,
and his garments, and his sons, and his sons' garments with him. Also
thou shalt take of the ram the fat and the rump, and the fat that
covereth the inwards, and the \emph{caul} above the liver, and the two
kidneys, and the fat that \emph{is} upon them, and the right shoulder;
for it \emph{is} a ram of consecration: And one loaf of bread, and one
cake of oiled bread, and one wafer out of the basket of the unleavened
bread that \emph{is} before the \textsc{Lord}: And thou shalt put all in
the hands of Aaron, and in the hands of his sons; and shalt wave them
\emph{for} a wave offering before the \textsc{Lord}. And thou shalt
receive them of their hands, and burn \emph{them} upon the altar for a
burnt offering, for a sweet savour before the \textsc{Lord}: it
\emph{is} an offering made by fire unto the \textsc{Lord}. And thou
shalt take the breast of the ram of Aaron's consecration, and wave it
\emph{for} a wave offering before the \textsc{Lord}: and it shall be thy
part. And thou shalt sanctify the breast of the wave offering, and the
shoulder of the heave offering, which is waved, and which is heaved up,
of the ram of the consecration, \emph{even} of \emph{that} which
\emph{is} for Aaron, and of \emph{that} which is for his sons: And it
shall be Aaron's and his sons' by a statute for ever from the children
of Israel: for it \emph{is} an heave offering: and it shall be an heave
offering from the children of Israel of the sacrifice of their peace
offerings, \emph{even} their heave offering unto the \textsc{Lord}.

And the holy garments of Aaron shall be his sons' after him, to be
anointed therein, and to be consecrated in them. \emph{And} that son
that is priest in his stead shall put them on seven days, when he cometh
into the tabernacle of the congregation to minister in the holy
\emph{place.}

And thou shalt take the ram of the consecration, and seethe his flesh in
the holy place. And Aaron and his sons shall eat the flesh of the ram,
and the bread that \emph{is} in the basket, \emph{by} the door of the
tabernacle of the congregation. And they shall eat those things
wherewith the atonement was made, to consecrate \emph{and} to sanctify
them: but a stranger shall not eat \emph{thereof,} because they
\emph{are} holy. And if ought of the flesh of the consecrations, or of
the bread, remain unto the morning, then thou shalt burn the remainder
with fire: it shall not be eaten, because it \emph{is} holy. And thus
shalt thou do unto Aaron, and to his sons, according to all
\emph{things} which I have commanded thee: seven days shalt thou
consecrate them. And thou shalt offer every day a bullock \emph{for} a
sin offering for atonement: and thou shalt cleanse the altar, when thou
hast made an atonement for it, and thou shalt anoint it, to sanctify it.
Seven days thou shalt make an atonement for the altar, and sanctify it;
and it shall be an altar most holy: whatsoever toucheth the altar shall
be holy.

Now this \emph{is that} which thou shalt offer upon the altar; two lambs
of the first year day by day continually. The one lamb thou shalt offer
in the morning; and the other lamb thou shalt offer at even: And with
the one lamb a tenth deal of flour mingled with the fourth part of an
hin of beaten oil; and the fourth part of an hin of wine \emph{for} a
drink offering. And the other lamb thou shalt offer at even, and shalt
do thereto according to the meat offering of the morning, and according
to the drink offering thereof, for a sweet savour, an offering made by
fire unto the \textsc{Lord}. \emph{This shall be} a continual burnt
offering throughout your generations \emph{at} the door of the
tabernacle of the congregation before the \textsc{Lord}: where I will
meet you, to speak there unto thee. And there I will meet with the
children of Israel, and \emph{the tabernacle} shall be sanctified by my
glory. And I will sanctify the tabernacle of the congregation, and the
altar: I will sanctify also both Aaron and his sons, to minister to me
in the priest's office.

And I will dwell among the children of Israel, and will be their God.
And they shall know that I \emph{am} the \textsc{Lord} their God, that
brought them forth out of the land of Egypt, that I may dwell among
them: I \emph{am} the \textsc{Lord} their God.

And thou shalt make an altar to burn incense upon: \emph{of} shittim
wood shalt thou make it. A cubit \emph{shall be} the length thereof, and
a cubit the breadth thereof; foursquare shall it be: and two cubits
\emph{shall be} the height thereof: the horns thereof \emph{shall be} of
the same. And thou shalt overlay it with pure gold, the top thereof, and
the sides thereof round about, and the horns thereof; and thou shalt
make unto it a crown of gold round about. And two golden rings shalt
thou make to it under the crown of it, by the two corners thereof, upon
the two sides of it shalt thou make \emph{it;} and they shall be for
places for the staves to bear it withal. And thou shalt make the staves
\emph{of} shittim wood, and overlay them with gold. And thou shalt put
it before the vail that \emph{is} by the ark of the testimony, before
the mercy seat that \emph{is} over the testimony, where I will meet with
thee. And Aaron shall burn thereon sweet incense every morning: when he
dresseth the lamps, he shall burn incense upon it. And when Aaron
lighteth the lamps at even, he shall burn incense upon it, a perpetual
incense before the \textsc{Lord} throughout your generations. Ye shall
offer no strange incense thereon, nor burnt sacrifice, nor meat
offering; neither shall ye pour drink offering thereon. And Aaron shall
make an atonement upon the horns of it once in a year with the blood of
the sin offering of atonements: once in the year shall he make atonement
upon it throughout your generations: it \emph{is} most holy unto the
\textsc{Lord}.

And the \textsc{Lord} spake unto Moses, saying, When thou takest the sum
of the children of Israel after their number, then shall they give every
man a ransom for his soul unto the \textsc{Lord}, when thou numberest
them; that there be no plague among them, when \emph{thou} numberest
them. This they shall give, every one that passeth among them that are
numbered, half a shekel after the shekel of the sanctuary: (a shekel
\emph{is} twenty gerahs:) an half shekel \emph{shall be} the offering of
the \textsc{Lord}. Every one that passeth among them that are numbered,
from twenty years old and above, shall give an offering unto the
\textsc{Lord}. The rich shall not give more, and the poor shall not give
less than half a shekel, when \emph{they} give an offering unto the
\textsc{Lord}, to make an atonement for your souls. And thou shalt take
the atonement money of the children of Israel, and shalt appoint it for
the service of the tabernacle of the congregation; that it may be a
memorial unto the children of Israel before the \textsc{Lord}, to make
an atonement for your souls.

And the \textsc{Lord} spake unto Moses, saying, Thou shalt also make a
laver \emph{of} brass, and his foot \emph{also of} brass, to wash
\emph{withal:} and thou shalt put it between the tabernacle of the
congregation and the altar, and thou shalt put water therein. For Aaron
and his sons shall wash their hands and their feet thereat: When they go
into the tabernacle of the congregation, they shall wash with water,
that they die not; or when they come near to the altar to minister, to
burn offering made by fire unto the \textsc{Lord}: So they shall wash
their hands and their feet, that they die not: and it shall be a statute
for ever to them, \emph{even} to him and to his seed throughout their
generations.

Moreover the \textsc{Lord} spake unto Moses, saying, Take thou also unto
thee principal spices, of pure myrrh five hundred \emph{shekels,} and of
sweet cinnamon half so much, \emph{even} two hundred and fifty
\emph{shekels,} and of sweet calamus two hundred and fifty \emph{shekels,
}And of cassia five hundred \emph{shekels,} after the shekel of the
sanctuary, and of oil olive an hin: And thou shalt make it an oil of
holy ointment, an ointment compound after the art of the apothecary: it
shall be an holy anointing oil. And thou shalt anoint the tabernacle of
the congregation therewith, and the ark of the testimony, And the table
and all his vessels, and the candlestick and his vessels, and the altar
of incense, And the altar of burnt offering with all his vessels, and
the laver and his foot. And thou shalt sanctify them, that they may be
most holy: whatsoever toucheth them shall be holy. And thou shalt anoint
Aaron and his sons, and consecrate them, that \emph{they} may minister
unto me in the priest's office. And thou shalt speak unto the children
of Israel, saying, This shall be an holy anointing oil unto me
throughout your generations. Upon man's flesh shall it not be poured,
neither shall ye make \emph{any other} like it, after the composition of
it: it \emph{is} holy, \emph{and} it shall be holy unto you. Whosoever
compoundeth \emph{any} like it, or whosoever putteth \emph{any} of it
upon a stranger, shall even be cut off from his people.

And the \textsc{Lord} said unto Moses, Take unto thee sweet spices,
stacte, and onycha, and galbanum; \emph{these} sweet spices with pure
frankincense: of each shall there be a like \emph{weight: }And thou
shalt make it a perfume, a confection after the art of the apothecary,
tempered together, pure \emph{and} holy: And thou shalt beat \emph{some}
of it very small, and put of it before the testimony in the tabernacle
of the congregation, where I will meet with thee: it shall be unto you
most holy. And \emph{as for} the perfume which thou shalt make, ye shall
not make to yourselves according to the composition thereof: it shall be
unto thee holy for the \textsc{Lord}. Whosoever shall make like unto
that, to smell thereto, shall even be cut off from his people.

And the \textsc{Lord} spake unto Moses, saying, See, I have called by
name Bezaleel the son of Uri, the son of Hur, of the tribe of Judah: And
I have filled him with the spirit of God, in wisdom, and in
understanding, and in knowledge, and in all manner of workmanship, To
devise cunning works, to work in gold, and in silver, and in brass, And
in cutting of stones, to set \emph{them,} and in carving of timber, to
work in all manner of workmanship. And I, behold, I have given with him
Aholiab, the son of Ahisamach, of the tribe of Dan: and in the hearts of
all that are wise hearted I have put wisdom, that they may make all that
I have commanded thee; The tabernacle of the congregation, and the ark
of the testimony, and the mercy seat that \emph{is} thereupon, and all
the furniture of the tabernacle, And the table and his furniture, and
the pure candlestick with all his furniture, and the altar of incense,
And the altar of burnt offering with all his furniture, and the laver
and his foot, And the cloths of service, and the holy garments for Aaron
the priest, and the garments of his sons, to minister in the priest's
office, And the anointing oil, and sweet incense for the holy
\emph{place:} according to all that I have commanded thee shall they do.

And the \textsc{Lord} spake unto Moses, saying, Speak thou also unto the
children of Israel, saying, Verily my sabbaths ye shall keep: for it is
a sign between me and you throughout your generations; that ye may know
that I \emph{am} the \textsc{Lord} that doth sanctify you. Ye shall keep
the sabbath therefore; for it \emph{is} holy unto you: every one that
defileth it shall surely be put to death: for whosoever doeth \emph{any}
work therein, that soul shall be cut off from among his people. Six days
may work be done; but in the seventh is the sabbath of rest, holy to the
\textsc{Lord}: whosoever doeth \emph{any} work in the sabbath day, he
shall surely be put to death. Wherefore the children of Israel shall
keep the sabbath, to observe the sabbath throughout their generations,
\emph{for} a perpetual covenant. It \emph{is} a sign between me and the
children of Israel for ever: for \emph{in} six days the \textsc{Lord}
made heaven and earth, and on the seventh day he rested, and was
refreshed.

And he gave unto Moses, when he had made an end of communing with him
upon mount Sinai, two tables of testimony, tables of stone, written with
the finger of God.

And when the people saw that Moses delayed to come down out of the
mount, the people gathered themselves together unto Aaron, and said unto
him, Up, make us gods, which shall go before us; for \emph{as for} this
Moses, the man that brought us up out of the land of Egypt, we wot not
what is become of him. And Aaron said unto them, Break off the golden
earrings, which \emph{are} in the ears of your wives, of your sons, and
of your daughters, and bring \emph{them} unto me. And all the people
brake off the golden earrings which \emph{were} in their ears, and
brought \emph{them} unto Aaron. And he received \emph{them} at their
hand, and fashioned it with a graving tool, after he had made it a
molten calf: and they said, These \emph{be} thy gods, O Israel, which
brought thee up out of the land of Egypt. And when Aaron saw \emph{it,}
he built an altar before it; and Aaron made proclamation, and said, To
morrow \emph{is} a feast to the \textsc{Lord}. And they rose up early on
the morrow, and offered burnt offerings, and brought peace offerings;
and the people sat down to eat and to drink, and rose up to play.

And the \textsc{Lord} said unto Moses, Go, get thee down; for thy
people, which thou broughtest out of the land of Egypt, have corrupted
\emph{themselves: }They have turned aside quickly out of the way which I
commanded them: they have made them a molten calf, and have worshipped
it, and have sacrificed thereunto, and said, These \emph{be} thy gods, O
Israel, which have brought thee up out of the land of Egypt. And the
\textsc{Lord} said unto Moses, I have seen this people, and, behold, it
\emph{is} a stiffnecked people: Now therefore let me alone, that my
wrath may wax hot against them, and that I may consume them: and I will
make of thee a great nation. And Moses besought the \textsc{Lord} his
God, and said, \textsc{Lord}, why doth thy wrath wax hot against thy
people, which thou hast brought forth out of the land of Egypt with
great power, and with a mighty hand? Wherefore should the Egyptians
speak, and say, For mischief did he bring them out, to slay them in the
mountains, and to consume them from the face of the earth? Turn from thy
fierce wrath, and repent of this evil against thy people. Remember
Abraham, Isaac, and Israel, thy servants, to whom thou swarest by thine
own self, and saidst unto them, I will multiply your seed as the stars
of heaven, and all this land that I have spoken of will I give unto your
seed, and they shall inherit \emph{it} for ever. And the \textsc{Lord}
repented of the evil which he thought to do unto his people.

And Moses turned, and went down from the mount, and the two tables of
the testimony \emph{were} in his hand: the tables \emph{were} written on
both their sides; on the one side and on the other \emph{were} they
written. And the tables \emph{were} the work of God, and the writing
\emph{was} the writing of God, graven upon the tables. And when Joshua
heard the noise of the people as they shouted, he said unto Moses,
\emph{There is} a noise of war in the camp. And he said, \emph{It is} not the
voice of \emph{them that} shout for mastery, neither \emph{is it} the
voice of \emph{them that} cry for being overcome: \emph{but} the noise
of \emph{them that} sing do I hear.

And it came to pass, as soon as he came nigh unto the camp, that he saw
the calf, and the dancing: and Moses' anger waxed hot, and he cast the
tables out of his hands, and brake them beneath the mount. And he took
the calf which they had made, and burnt \emph{it} in the fire, and
ground \emph{it} to powder, and strawed \emph{it} upon the water, and
made the children of Israel drink \emph{of it. }And Moses said unto
Aaron, What did this people unto thee, that thou hast brought so great
a sin upon them? And Aaron said, Let not the anger of my lord wax hot:
thou knowest the people, that they are \emph{set} on mischief. For they
said unto me, Make us gods, which shall go before us: for \emph{as for}
this Moses, the man that brought us up out of the land of Egypt, we wot
not what is become of him. And I said unto them, Whosoever hath any
gold, let them break \emph{it} off. So they gave \emph{it} me: then I
cast it into the fire, and there came out this calf.

And when Moses saw that the people \emph{were} naked; (for Aaron had
made them naked unto \emph{their} shame among their enemies:) Then Moses
stood in the gate of the camp, and said, Who \emph{is} on the
\textsc{Lord}'S side? \emph{let him come} unto me. And all the sons of
Levi gathered themselves together unto him. And he said unto them, Thus
saith the \textsc{Lord} God of Israel, Put every man his sword by his
side, \emph{and} go in and out from gate to gate throughout the camp,
and slay every man his brother, and every man his companion, and every
man his neighbour. And the children of Levi did according to the word of
Moses: and there fell of the people that day about three thousand men.
For Moses had said, Consecrate yourselves to day to the \textsc{Lord},
even every man upon his son, and upon his brother; that he may bestow
upon you a blessing this day.

And it came to pass on the morrow, that Moses said unto the people, Ye
have sinned a great sin: and now I will go up unto the \textsc{Lord};
peradventure I shall make an atonement for your sin. And Moses returned
unto the \textsc{Lord}, and said, Oh, this people have sinned a great
sin, and have made them gods of gold. Yet now, if thou wilt forgive
their sin--; and if not, blot me, I pray thee, out of thy book which
thou hast written. And the \textsc{Lord} said unto Moses, Whosoever hath
sinned against me, him will I blot out of my book. Therefore now go,
lead the people unto \emph{the place} of which I have spoken unto thee:
behold, mine Angel shall go before thee: nevertheless in the day when I
visit I will visit their sin upon them. And the \textsc{Lord} plagued
the people, because they made the calf, which Aaron made.

And the \textsc{Lord} said unto Moses, Depart, \emph{and} go up hence,
thou and the people which thou hast brought up out of the land of Egypt,
unto the land which I sware unto Abraham, to Isaac, and to Jacob,
saying, Unto thy seed will I give it: And I will send an angel before
thee; and I will drive out the Canaanite, the Amorite, and the Hittite,
and the Perizzite, the Hivite, and the Jebusite: Unto a land flowing
with milk and honey: for I will not go up in the midst of thee; for thou
\emph{art} a stiffnecked people: lest I consume thee in the way.

And when the people heard these evil tidings, they mourned: and no man
did put on him his ornaments. For the \textsc{Lord} had said unto Moses,
Say unto the children of Israel, Ye \emph{are} a stiffnecked people: I
will come up into the midst of thee in a moment, and consume thee:
therefore now put off thy ornaments from thee, that I may know what to
do unto thee. And the children of Israel stripped themselves of their
ornaments by the mount Horeb. And Moses took the tabernacle, and pitched
it without the camp, afar off from the camp, and called it the
Tabernacle of the congregation. And it came to pass, \emph{that} every
one which sought the \textsc{Lord} went out unto the tabernacle of the
congregation, which \emph{was} without the camp. And it came to pass,
when Moses went out unto the tabernacle, \emph{that} all the people rose
up, and stood every man \emph{at} his tent door, and looked after Moses,
until he was gone into the tabernacle. And it came to pass, as Moses
entered into the tabernacle, the cloudy pillar descended, and stood
\emph{at} the door of the tabernacle, and \emph{the} \textsc{Lord}
talked with Moses. And all the people saw the cloudy pillar stand
\emph{at} the tabernacle door: and all the people rose up and
worshipped, every man \emph{in} his tent door. And the \textsc{Lord}
spake unto Moses face to face, as a man speaketh unto his friend. And he
turned again into the camp: but his servant Joshua, the son of Nun, a
young man, departed not out of the tabernacle.

And Moses said unto the \textsc{Lord}, See, thou sayest unto me, Bring
up this people: and thou hast not let me know whom thou wilt send with
me. Yet thou hast said, I know thee by name, and thou hast also found
grace in my sight. Now therefore, I pray thee, if I have found grace in
thy sight, shew me now thy way, that I may know thee, that I may find
grace in thy sight: and consider that this nation \emph{is} thy people.
And he said, My presence shall go \emph{with thee,} and I will give thee
rest. And he said unto him, If thy presence go not \emph{with me,} carry
us not up hence. For wherein shall it be known here that I and thy
people have found grace in thy sight? \emph{is it} not in that thou
goest with us? so shall we be separated, I and thy people, from all the
people that \emph{are} upon the face of the earth. And the \textsc{Lord}
said unto Moses, I will do this thing also that thou hast spoken: for
thou hast found grace in my sight, and I know thee by name. And he said,
I beseech thee, shew me thy glory. And he said, I will make all my
goodness pass before thee, and I will proclaim the name of the
\textsc{Lord} before thee; and will be gracious to whom I will be
gracious, and will shew mercy on whom I will shew mercy. And he said,
Thou canst not see my face: for there shall no man see me, and live. And
the \textsc{Lord} said, Behold, \emph{there is} a place by me, and thou
shalt stand upon a rock: And it shall come to pass, while my glory
passeth by, that I will put thee in a clift of the rock, and will cover
thee with my hand while I pass by: And I will take away mine hand, and
thou shalt see my back parts: but my face shall not be seen.

And the \textsc{Lord} said unto Moses, Hew thee two tables of stone like
unto the first: and I will write upon \emph{these} tables the words that
were in the first tables, which thou brakest. And be ready in the
morning, and come up in the morning unto mount Sinai, and present
thyself there to me in the top of the mount. And no man shall come up
with thee, neither let any man be seen throughout all the mount; neither
let the flocks nor herds feed before that mount.

And he hewed two tables of stone like unto the first; and Moses rose up
early in the morning, and went up unto mount Sinai, as the \textsc{Lord}
had commanded him, and took in his hand the two tables of stone. And the
\textsc{Lord} descended in the cloud, and stood with him there, and
proclaimed the name of the \textsc{Lord}. And the \textsc{Lord} passed
by before him, and proclaimed, The \textsc{Lord}, The \textsc{Lord} God,
merciful and gracious, longsuffering, and abundant in goodness and
truth, Keeping mercy for thousands, forgiving iniquity and transgression
and sin, and that will by no means clear \emph{the guilty;} visiting the
iniquity of the fathers upon the children, and upon the children's
children, unto the third and to the fourth \emph{generation. }And Moses
made haste, and bowed his head toward the earth, and worshipped. And he
said, If now I have found grace in thy sight, O Lord, let my Lord, I
pray thee, go among us; for it \emph{is} a stiffnecked people; and
pardon our iniquity and our sin, and take us for thine inheritance.

And he said, Behold, I make a covenant: before all thy people I will do
marvels, such as have not been done in all the earth, nor in any nation:
and all the people among which thou \emph{art} shall see the work of the
\textsc{Lord}: for it \emph{is} a terrible thing that I will do with
thee. Observe thou that which I command thee this day: behold, I drive
out before thee the Amorite, and the Canaanite, and the Hittite, and the
Perizzite, and the Hivite, and the Jebusite. Take heed to thyself, lest
thou make a covenant with the inhabitants of the land whither thou
goest, lest it be for a snare in the midst of thee: But ye shall destroy
their altars, break their images, and cut down their groves: For thou
shalt worship no other god: for the \textsc{Lord}, whose name \emph{is}
Jealous, \emph{is} a jealous God: Lest thou make a covenant with the
inhabitants of the land, and they go a whoring after their gods, and do
sacrifice unto their gods, and \emph{one} call thee, and thou eat of his
sacrifice; And thou take of their daughters unto thy sons, and their
daughters go a whoring after their gods, and make thy sons go a whoring
after their gods. Thou shalt make thee no molten gods.

The feast of unleavened bread shalt thou keep. Seven days thou shalt eat
unleavened bread, as I commanded thee, in the time of the month Abib:
for in the month Abib thou camest out from Egypt. All that openeth the
matrix \emph{is} mine; and every firstling among thy cattle,
\emph{whether} ox or sheep, \emph{that is male. }But the firstling of an
ass thou shalt redeem with a lamb: and if thou redeem \emph{him} not,
then shalt thou break his neck. All the firstborn of thy sons thou shalt
redeem. And none shall appear before me empty.

Six days thou shalt work, but on the seventh day thou shalt rest: in
earing time and in harvest thou shalt rest.

And thou shalt observe the feast of weeks, of the firstfruits of wheat
harvest, and the feast of ingathering at the year's end.

Thrice in the year shall all your men children appear before the Lord
GOD, the God of Israel. For I will cast out the nations before thee, and
enlarge thy borders: neither shall any man desire thy land, when thou
shalt go up to appear before the \textsc{Lord} thy God thrice in the
year. Thou shalt not offer the blood of my sacrifice with leaven;
neither shall the sacrifice of the feast of the passover be left unto
the morning. The first of the firstfruits of thy land thou shalt bring
unto the house of the \textsc{Lord} thy God. Thou shalt not seethe a kid
in his mother's milk. And the \textsc{Lord} said unto Moses, Write thou
these words: for after the tenor of these words I have made a covenant
with thee and with Israel. And he was there with the \textsc{Lord} forty
days and forty nights; he did neither eat bread, nor drink water. And he
wrote upon the tables the words of the covenant, the ten commandments.

And it came to pass, when Moses came down from mount Sinai with the two
tables of testimony in Moses' hand, when he came down from the mount,
that Moses wist not that the skin of his face shone while he talked with
him. And when Aaron and all the children of Israel saw Moses, behold,
the skin of his face shone; and they were afraid to come nigh him. And
Moses called unto them; and Aaron and all the rulers of the congregation
returned unto him: and Moses talked with them. And afterward all the
children of Israel came nigh: and he gave them in commandment all that
the \textsc{Lord} had spoken with him in mount Sinai. And \emph{till}
Moses had done speaking with them, he put a vail on his face. But when
Moses went in before the \textsc{Lord} to speak with him, he took the
vail off, until he came out. And he came out, and spake unto the
children of Israel \emph{that} which he was commanded. And the children
of Israel saw the face of Moses, that the skin of Moses' face shone: and
Moses put the vail upon his face again, until he went in to speak with
him.

And Moses gathered all the congregation of the children of Israel
together, and said unto them, These \emph{are} the words which the
\textsc{Lord} hath commanded, that \emph{ye} should do them. Six days
shall work be done, but on the seventh day there shall be to you an holy
day, a sabbath of rest to the \textsc{Lord}: whosoever doeth work
therein shall be put to death. Ye shall kindle no fire throughout your
habitations upon the sabbath day.

And Moses spake unto all the congregation of the children of Israel,
saying, This \emph{is} the thing which the \textsc{Lord} commanded,
saying, Take ye from among you an offering unto the \textsc{Lord}:
whosoever \emph{is} of a willing heart, let him bring it, an offering of
the \textsc{Lord}; gold, and silver, and brass, And blue, and purple,
and scarlet, and fine linen, and goats' \emph{hair, }And rams' skins
dyed red, and badgers' skins, and shittim wood, And oil for the light,
and spices for anointing oil, and for the sweet incense, And onyx
stones, and stones to be set for the ephod, and for the breastplate. And
every wise hearted among you shall come, and make all that the
\textsc{Lord} hath commanded; The tabernacle, his tent, and his
covering, his taches, and his boards, his bars, his pillars, and his
sockets, The ark, and the staves thereof, \emph{with} the mercy seat,
and the vail of the covering, The table, and his staves, and all his
vessels, and the shewbread, The candlestick also for the light, and his
furniture, and his lamps, with the oil for the light, And the incense
altar, and his staves, and the anointing oil, and the sweet incense, and
the hanging for the door at the entering in of the tabernacle, The altar
of burnt offering, with his brasen grate, his staves, and all his
vessels, the laver and his foot, The hangings of the court, his pillars,
and their sockets, and the hanging for the door of the court, The pins
of the tabernacle, and the pins of the court, and their cords, The
cloths of service, to do service in the holy \emph{place,} the holy
garments for Aaron the priest, and the garments of his sons, to minister
in the priest's office.

And all the congregation of the children of Israel departed from the
presence of Moses. And they came, every one whose heart stirred him up,
and every one whom his spirit made willing, \emph{and} they brought the
\textsc{Lord}'S offering to the work of the tabernacle of the
congregation, and for all his service, and for the holy garments. And
they came, both men and women, as many as were willing hearted,
\emph{and} brought bracelets, and earrings, and rings, and tablets, all
jewels of gold: and every man that offered \emph{offered} an offering of
gold unto the \textsc{Lord}. And every man, with whom was found blue,
and purple, and scarlet, and fine linen, and goats' \emph{hair,} and red
skins of rams, and badgers' skins, brought \emph{them. }Every one that
did offer an offering of silver and brass brought the \textsc{Lord}'S
offering: and every man, with whom was found shittim wood for any work
of the service, brought \emph{it. }And all the women that were wise
hearted did spin with their hands, and brought that which they had spun,
\emph{both} of blue, and of purple, \emph{and} of scarlet, and of fine
linen. And all the women whose heart stirred them up in wisdom spun
goats' \emph{hair. }And the rulers brought onyx stones, and stones to be
set, for the ephod, and for the breastplate; And spice, and oil for the
light, and for the anointing oil, and for the sweet incense. The
children of Israel brought a willing offering unto the \textsc{Lord},
every man and woman, whose heart made them willing to bring for all
manner of work, which the \textsc{Lord} had commanded to be made by the
hand of Moses.

And Moses said unto the children of Israel, See, the \textsc{Lord} hath
called by name Bezaleel the son of Uri, the son of Hur, of the tribe of
Judah; And he hath filled him with the spirit of God, in wisdom, in
understanding, and in knowledge, and in all manner of workmanship; And
to devise curious works, to work in gold, and in silver, and in brass,
And in the cutting of stones, to set \emph{them,} and in carving of
wood, to make any manner of cunning work. And he hath put in his heart
that he may teach, \emph{both} he, and Aholiab, the son of Ahisamach, of
the tribe of Dan. Them hath he filled with wisdom of heart, to work all
manner of work, of the engraver, and of the cunning workman, and of the
embroiderer, in blue, and in purple, in scarlet, and in fine linen, and
of the weaver, \emph{even} of them that do any work, and of those that
devise cunning work.

Then wrought Bezaleel and Aholiab, and every wise hearted man, in whom
the \textsc{Lord} put wisdom and understanding to know how to work all
manner of work for the service of the sanctuary, according to all that
the \textsc{Lord} had commanded. And Moses called Bezaleel and Aholiab,
and every wise hearted man, in whose heart the \textsc{Lord} had put
wisdom, \emph{even} every one whose heart stirred him up to come unto
the work to do it: And they received of Moses all the offering, which
the children of Israel had brought for the work of the service of the
sanctuary, to make it \emph{withal.} And they brought yet unto him free
offerings every morning. And all the wise men, that wrought all the work
of the sanctuary, came every man from his work which they made;

And they spake unto Moses, saying, The people bring much more than
enough for the service of the work, which the \textsc{Lord} commanded to
make. And Moses gave commandment, and they caused it to be proclaimed
throughout the camp, saying, Let neither man nor woman make any more
work for the offering of the sanctuary. So the people were restrained
from bringing. For the stuff they had was sufficient for all the work to
make it, and too much.

And every wise hearted man among them that wrought the work of the
tabernacle made ten curtains \emph{of} fine twined linen, and blue, and
purple, and scarlet: \emph{with} cherubims of cunning work made he them.
The length of one curtain \emph{was} twenty and eight cubits, and the
breadth of one curtain four cubits: the curtains \emph{were} all of one
size. And he coupled the five curtains one unto another: and \emph{the
other} five curtains he coupled one unto another. And he made loops of
blue on the edge of one curtain from the selvedge in the coupling:
likewise he made in the uttermost side of \emph{another} curtain, in the
coupling of the second. Fifty loops made he in one curtain, and fifty
loops made he in the edge of the curtain which \emph{was} in the
coupling of the second: the loops held one \emph{curtain} to another.
And he made fifty taches of gold, and coupled the curtains one unto
another with the taches: so it became one tabernacle.

And he made curtains \emph{of} goats' \emph{hair} for the tent over the
tabernacle: eleven curtains he made them. The length of one curtain
\emph{was} thirty cubits, and four cubits \emph{was} the breadth of one
curtain: the eleven curtains \emph{were} of one size. And he coupled
five curtains by themselves, and six curtains by themselves. And he made
fifty loops upon the uttermost edge of the curtain in the coupling, and
fifty loops made he upon the edge of the curtain which coupleth the
second. And he made fifty taches \emph{of} brass to couple the tent
together, that it might be one. And he made a covering for the tent
\emph{of} rams' skins dyed red, and a covering \emph{of} badgers' skins
above \emph{that.}

And he made boards for the tabernacle \emph{of} shittim wood, standing
up. The length of a board \emph{was} ten cubits, and the breadth of a
board one cubit and a half. One board had two tenons, equally distant
one from another: thus did he make for all the boards of the tabernacle.
And he made boards for the tabernacle; twenty boards for the south side
southward: And forty sockets of silver he made under the twenty boards;
two sockets under one board for his two tenons, and two sockets under
another board for his two tenons. And for the other side of the
tabernacle, \emph{which is} toward the north corner, he made twenty
boards, And their forty sockets of silver; two sockets under one board,
and two sockets under another board. And for the sides of the tabernacle
westward he made six boards. And two boards made he for the corners of
the tabernacle in the two sides. And they were coupled beneath, and
coupled together at the head thereof, to one ring: thus he did to both
of them in both the corners. And there were eight boards; and their
sockets \emph{were} sixteen sockets of silver, under every board two
sockets.

And he made bars of shittim wood; five for the boards of the one side of
the tabernacle, And five bars for the boards of the other side of the
tabernacle, and five bars for the boards of the tabernacle for the sides
westward. And he made the middle bar to shoot through the boards from
the one end to the other. And he overlaid the boards with gold, and made
their rings \emph{of} gold \emph{to be} places for the bars, and overlaid the
bars with gold.

And he made a vail \emph{of} blue, and purple, and scarlet, and fine
twined linen: \emph{with} cherubims made he it of cunning work. And he
made thereunto four pillars \emph{of} shittim \emph{wood,} and overlaid
them with gold: their hooks \emph{were of} gold; and he cast for them
four sockets of silver.

And he made an hanging for the tabernacle door \emph{of} blue, and
purple, and scarlet, and fine twined linen, of needlework; And the five
pillars of it with their hooks: and he overlaid their chapiters and
their fillets with gold: but their five sockets \emph{were of} brass.

And Bezaleel made the ark \emph{of} shittim wood: two cubits and a half
\emph{was} the length of it, and a cubit and a half the breadth of it,
and a cubit and a half the height of it: And he overlaid it with pure
gold within and without, and made a crown of gold to it round about. And
he cast for it four rings of gold, \emph{to be set} by the four corners
of it; even two rings upon the one side of it, and two rings upon the
other side of it. And he made staves \emph{of} shittim wood, and
overlaid them with gold. And he put the staves into the rings by the
sides of the ark, to bear the ark.

And he made the mercy seat \emph{of} pure gold: two cubits and a half
\emph{was} the length thereof, and one cubit and a half the breadth
thereof. And he made two cherubims \emph{of} gold, beaten out of one
piece made he them, on the two ends of the mercy seat; One cherub on the
end on this side, and another cherub on the \emph{other} end on that
side: out of the mercy seat made he the cherubims on the two ends
thereof. And the cherubims spread out \emph{their} wings on high,
\emph{and} covered with their wings over the mercy seat, with their
faces one to another; \emph{even }to the mercy seatward were the faces
of the cherubims.

And he made the table of shittim wood: two cubits \emph{was} the length
thereof, and a cubit the breadth thereof, and a cubit and a half the
height thereof: And he overlaid it with pure gold, and made thereunto a
crown of gold round about. Also he made thereunto a border of an
handbreadth round about; and made a crown of gold for the border thereof
round about. And he cast for it four rings of gold, and put the rings
upon the four corners that \emph{were} in the four feet thereof. Over
against the border were the rings, the places for the staves to bear the
table. And he made the staves \emph{of} shittim wood, and overlaid them
with gold, to bear the table. And he made the vessels which \emph{were}
upon the table, his dishes, and his spoons, and his bowls, and his
covers to cover withal, \emph{of} pure gold.

And he made the candlestick \emph{of} pure gold: \emph{of} beaten work
made he the candlestick; his shaft, and his branch, his bowls, his
knops, and his flowers, were of the same: And six branches going out of
the sides thereof; three branches of the candlestick out of the one side
thereof, and three branches of the candlestick out of the other side
thereof: Three bowls made after the fashion of almonds in one branch, a
knop and a flower; and three bowls made like almonds in another branch,
a knop and a flower: so throughout the six branches going out of the
candlestick. And in the candlestick \emph{were} four bowls made like
almonds, his knops, and his flowers: And a knop under two branches of
the same, and a knop under two branches of the same, and a knop under
two branches of the same, according to the six branches going out of it.
Their knops and their branches were of the same: all of it \emph{was}
one beaten work \emph{of} pure gold. And he made his seven lamps, and
his snuffers, and his snuffdishes, \emph{of} pure gold. \emph{Of} a
talent of pure gold made he it, and all the vessels thereof.

And he made the incense altar \emph{of} shittim wood: the length of it
\emph{was} a cubit, and the breadth of it a cubit; it \emph{was}
foursquare; and two cubits \emph{was} the height of it; the horns
thereof were of the same. And he overlaid it with pure gold, \emph{both}
the top of it, and the sides thereof round about, and the horns of it:
also he made unto it a crown of gold round about. And he made two rings
of gold for it under the crown thereof, by the two corners of it, upon
the two sides thereof, to be places for the staves to bear it withal.
And he made the staves \emph{of} shittim wood, and overlaid them with
gold.

And he made the holy anointing oil, and the pure incense of sweet
spices, according to the work of the apothecary.

And he made the altar of burnt offering \emph{of} shittim wood: five
cubits \emph{was} the length thereof, and five cubits the breadth
thereof; \emph{it was} foursquare; and three cubits the height thereof.
And he made the horns thereof on the four corners of it; the horns
thereof were of the same: and he overlaid it with brass. And he made all
the vessels of the altar, the pots, and the shovels, and the basons,
\emph{and} the fleshhooks, and the firepans: all the vessels thereof
made he \emph{of} brass. And he made for the altar a brasen grate of
network under the compass thereof beneath unto the midst of it. And he
cast four rings for the four ends of the grate of brass, \emph{to be}
places for the staves. And he made the staves \emph{of} shittim wood,
and overlaid them with brass. And he put the staves into the rings on
the sides of the altar, to bear it withal; he made the altar hollow with
boards.

And he made the laver \emph{of} brass, and the foot of it \emph{of}
brass, of the lookingglasses of \emph{the women} assembling, which
assembled \emph{at} the door of the tabernacle of the congregation.

And he made the court: on the south side southward the hangings of the
court \emph{were of} fine twined linen, an hundred cubits: Their pillars
\emph{were} twenty, and their brasen sockets twenty; the hooks of the
pillars and their fillets \emph{were of} silver. And for the north side
\emph{the hangings were} an hundred cubits, their pillars \emph{were}
twenty, and their sockets of brass twenty; the hooks of the pillars and
their fillets \emph{of} silver. And for the west side \emph{were}
hangings of fifty cubits, their pillars ten, and their sockets ten; the
hooks of the pillars and their fillets \emph{of} silver. And for the
east side eastward fifty cubits. The hangings of the one side \emph{of
the gate were} fifteen cubits; their pillars three, and their sockets
three. And for the other side of the court gate, on this hand and that
hand, \emph{were} hangings of fifteen cubits; their pillars three, and
their sockets three. All the hangings of the court round about
\emph{were} of fine twined linen. And the sockets for the pillars
\emph{were of} brass; the hooks of the pillars and their fillets
\emph{of} silver; and the overlaying of their chapiters \emph{of}
silver; and all the pillars of the court \emph{were} filleted with
silver. And the hanging for the gate of the court \emph{was} needlework,
\emph{of} blue, and purple, and scarlet, and fine twined linen: and
twenty cubits \emph{was} the length, and the height in the breadth
\emph{was} five cubits, answerable to the hangings of the court. And
their pillars \emph{were} four, and their sockets \emph{of} brass four;
their hooks \emph{of} silver, and the overlaying of their chapiters and
their fillets \emph{of} silver. And all the pins of the tabernacle, and
of the court round about, \emph{were of} brass.

This is the sum of the tabernacle, \emph{even} of the tabernacle of
testimony, as it was counted, according to the commandment of Moses,
\emph{for} the service of the Levites, by the hand of Ithamar, son to
Aaron the priest. And Bezaleel the son of Uri, the son of Hur, of the
tribe of Judah, made all that the \textsc{Lord} commanded Moses. And
with him \emph{was} Aholiab, son of Ahisamach, of the tribe of Dan, an
engraver, and a cunning workman, and an embroiderer in blue, and in
purple, and in scarlet, and fine linen. All the gold that was occupied
for the work in all the work of the holy \emph{place,} even the gold of
the offering, was twenty and nine talents, and seven hundred and thirty
shekels, after the shekel of the sanctuary. And the silver of them that
were numbered of the congregation \emph{was} an hundred talents, and a
thousand seven hundred and threescore and fifteen shekels, after the
shekel of the sanctuary: A bekah for every man, \emph{that is,} half a
shekel, after the shekel of the sanctuary, for every one that went to be
numbered, from twenty years old and upward, for six hundred thousand and
three thousand and five hundred and fifty \emph{men. }And of the hundred
talents of silver were cast the sockets of the sanctuary, and the
sockets of the vail; an hundred sockets of the hundred talents, a talent
for a socket. And of the thousand seven hundred seventy and five
\emph{shekels} he made hooks for the pillars, and overlaid their
chapiters, and filleted them. And the brass of the offering \emph{was}
seventy talents, and two thousand and four hundred shekels. And
therewith he made the sockets to the door of the tabernacle of the
congregation, and the brasen altar, and the brasen grate for it, and all
the vessels of the altar, And the sockets of the court round about, and
the sockets of the court gate, and all the pins of the tabernacle, and
all the pins of the court round about.

And of the blue, and purple, and scarlet, they made cloths of service,
to do service in the holy \emph{place,} and made the holy garments for
Aaron; as the \textsc{Lord} commanded Moses. And he made the ephod
\emph{of} gold, blue, and purple, and scarlet, and fine twined linen.
And they did beat the gold into thin plates, and cut \emph{it into}
wires, to work \emph{it} in the blue, and in the purple, and in the
scarlet, and in the fine linen, \emph{with} cunning work. They made
shoulderpieces for it, to couple \emph{it} together: by the two edges
was it coupled together. And the curious girdle of his ephod, that
\emph{was} upon it, \emph{was} of the same, according to the work
thereof; \emph{of} gold, blue, and purple, and scarlet, and fine twined
linen; as the \textsc{Lord} commanded Moses.

And they wrought onyx stones inclosed in ouches of gold, graven, as
signets are graven, with the names of the children of Israel. And he put
them on the shoulders of the ephod, \emph{that they should be} stones for a
memorial to the children of Israel; as the \textsc{Lord} commanded
Moses.

And he made the breastplate \emph{of} cunning work, like the work of the
ephod; \emph{of} gold, blue, and purple, and scarlet, and fine twined
linen. It was foursquare; they made the breastplate double: a span
\emph{was} the length thereof, and a span the breadth thereof,
\emph{being} doubled. And they set in it four rows of stones: \emph{the
first} row \emph{was} a sardius, a topaz, and a carbuncle: this
\emph{was} the first row. And the second row, an emerald, a sapphire,
and a diamond. And the third row, a ligure, an agate, and an amethyst.
And the fourth row, a beryl, an onyx, and a jasper: \emph{they were}
inclosed in ouches of gold in their inclosings. And the stones
\emph{were} according to the names of the children of Israel, twelve,
according to their names, \emph{like} the engravings of a signet, every
one with his name, according to the twelve tribes. And they made upon
the breastplate chains at the ends, \emph{of} wreathen work \emph{of}
pure gold. And they made two ouches \emph{of} gold, and two gold rings;
and put the two rings in the two ends of the breastplate. And they put
the two wreathen chains of gold in the two rings on the ends of the
breastplate. And the two ends of the two wreathen chains they fastened
in the two ouches, and put them on the shoulderpieces of the ephod,
before it. And they made two rings of gold, and put \emph{them} on the
two ends of the breastplate, upon the border of it, which \emph{was} on
the side of the ephod inward. And they made two \emph{other} golden
rings, and put them on the two sides of the ephod underneath, toward the
forepart of it, over against the \emph{other} coupling thereof, above
the curious girdle of the ephod. And they did bind the breastplate by
his rings unto the rings of the ephod with a lace of blue, that it might
be above the curious girdle of the ephod, and that the breastplate might
not be loosed from the ephod; as the \textsc{Lord} commanded Moses.

And he made the robe of the ephod \emph{of} woven work, all \emph{of}
blue. And \emph{there was} an hole in the midst of the robe, as the hole
of an habergeon, \emph{with} a band round about the hole, that it should
not rend. And they made upon the hems of the robe pomegranates \emph{of}
blue, and purple, and scarlet, \emph{and} twined \emph{linen. }And they made
bells \emph{of} pure gold, and put the bells between the pomegranates
upon the hem of the robe, round about between the pomegranates; A bell
and a pomegranate, a bell and a pomegranate, round about the hem of the
robe to minister \emph{in;} as the \textsc{Lord} commanded Moses.

And they made coats \emph{of} fine linen \emph{of} woven work for Aaron,
and for his sons, And a mitre \emph{of} fine linen, and goodly bonnets
\emph{of} fine linen, and linen breeches \emph{of} fine twined linen,
And a girdle \emph{of} fine twined linen, and blue, and purple, and
scarlet, \emph{of} needlework; as the \textsc{Lord} commanded Moses.

And they made the plate of the holy crown \emph{of} pure gold, and wrote
upon it a writing, \emph{like to} the engravings of a signet, HOLINESS
TO THE \textsc{Lord}. And they tied unto it a lace of blue, to fasten
\emph{it} on high upon the mitre; as the \textsc{Lord} commanded Moses.

Thus was all the work of the tabernacle of the tent of the congregation
finished: and the children of Israel did according to all that the
\textsc{Lord} commanded Moses, so did they.

And they brought the tabernacle unto Moses, the tent, and all his
furniture, his taches, his boards, his bars, and his pillars, and his
sockets, And the covering of rams' skins dyed red, and the covering of
badgers' skins, and the vail of the covering, The ark of the testimony,
and the staves thereof, and the mercy seat, The table, \emph{and} all
the vessels thereof, and the shewbread, The pure candlestick,
\emph{with} the lamps thereof, \emph{even with} the lamps to be set in
order, and all the vessels thereof, and the oil for light, And the
golden altar, and the anointing oil, and the sweet incense, and the
hanging for the tabernacle door, The brasen altar, and his grate of
brass, his staves, and all his vessels, the laver and his foot, The
hangings of the court, his pillars, and his sockets, and the hanging for
the court gate, his cords, and his pins, and all the vessels of the
service of the tabernacle, for the tent of the congregation, The cloths
of service to do service in the holy \emph{place,} and the holy garments
for Aaron the priest, and his sons' garments, to minister in the
priest's office. According to all that the \textsc{Lord} commanded
Moses, so the children of Israel made all the work. And Moses did look
upon all the work, and, behold, they have done it as the \textsc{Lord}
had commanded, even so had they done it: and Moses blessed them.

And the \textsc{Lord} spake unto Moses, saying, On the first day of the
first month shalt thou set up the tabernacle of the tent of the
congregation. And thou shalt put therein the ark of the testimony, and
cover the ark with the vail. And thou shalt bring in the table, and set
in order the things that are to be set in order upon it; and thou shalt
bring in the candlestick, and light the lamps thereof. And thou shalt
set the altar of gold for the incense before the ark of the testimony,
and put the hanging of the door to the tabernacle. And thou shalt set
the altar of the burnt offering before the door of the tabernacle of the
tent of the congregation. And thou shalt set the laver between the tent
of the congregation and the altar, and shalt put water therein. And thou
shalt set up the court round about, and hang up the hanging at the court
gate. And thou shalt take the anointing oil, and anoint the tabernacle,
and all that \emph{is} therein, and shalt hallow it, and all the vessels
thereof: and it shall be holy. And thou shalt anoint the altar of the
burnt offering, and all his vessels, and sanctify the altar: and it
shall be an altar most holy. And thou shalt anoint the laver and his
foot, and sanctify it. And thou shalt bring Aaron and his sons unto the
door of the tabernacle of the congregation, and wash them with water.
And thou shalt put upon Aaron the holy garments, and anoint him, and
sanctify him; that he may minister unto me in the priest's office. And
thou shalt bring his sons, and clothe them with coats: And thou shalt
anoint them, as thou didst anoint their father, that they may minister
unto me in the priest's office: for their anointing shall surely be an
everlasting priesthood throughout their generations. Thus did Moses:
according to all that the \textsc{Lord} commanded him, so did he.

And it came to pass in the first month in the second year, on the first
\emph{day} of the month, \emph{that} the tabernacle was reared up. And
Moses reared up the tabernacle, and fastened his sockets, and set up the
boards thereof, and put in the bars thereof, and reared up his pillars.
And he spread abroad the tent over the tabernacle, and put the covering
of the tent above upon it; as the \textsc{Lord} commanded Moses.

And he took and put the testimony into the ark, and set the staves on
the ark, and put the mercy seat above upon the ark: And he brought the
ark into the tabernacle, and set up the vail of the covering, and
covered the ark of the testimony; as the \textsc{Lord} commanded Moses.

And he put the table in the tent of the congregation, upon the side of
the tabernacle northward, without the vail. And he set the bread in
order upon it before the \textsc{Lord}; as the \textsc{Lord} had
commanded Moses.

And he put the candlestick in the tent of the congregation, over against
the table, on the side of the tabernacle southward. And he lighted the
lamps before the \textsc{Lord}; as the \textsc{Lord} commanded Moses.

And he put the golden altar in the tent of the congregation before the
vail: And he burnt sweet incense thereon; as the \textsc{Lord} commanded
Moses.

And he set up the hanging \emph{at} the door of the tabernacle. And he
put the altar of burnt offering \emph{by} the door of the tabernacle of
the tent of the congregation, and offered upon it the burnt offering and
the meat offering; as the \textsc{Lord} commanded Moses.

And he set the laver between the tent of the congregation and the altar,
and put water there, to wash \emph{withal. }And Moses and Aaron and his
sons washed their hands and their feet thereat: When they went into the
tent of the congregation, and when they came near unto the altar, they
washed; as the \textsc{Lord} commanded Moses. And he reared up the court
round about the tabernacle and the altar, and set up the hanging of the
court gate. So Moses finished the work.

Then a cloud covered the tent of the congregation, and the glory of the
\textsc{Lord} filled the tabernacle. And Moses was not able to enter
into the tent of the congregation, because the cloud abode thereon, and
the glory of the \textsc{Lord} filled the tabernacle. And when the cloud
was taken up from over the tabernacle, the children of Israel went
onward in all their journeys: But if the cloud were not taken up, then
they journeyed not till the day that it was taken up. For the cloud of
the \textsc{Lord} \emph{was} upon the tabernacle by day, and fire was on
it by night, in the sight of all the house of Israel, throughout all
their journeys.
\end{document}