\documentclass[11pt,letterpaper,oneside]{memoir}
\usepackage{charter}

% Set the typeblock size to ensure about 65 characters per line (with Charter 11pt, that's 
% 26 picas).
\settypeblocksize{8.5in}{26pc}{*}
% assuming room for a 3-hole punch (0.625in on the spine)
%\setlrmargins{1.625in}{*}{*}
\setlrmargins{12.5pc}{*}{*}
\setulmargins{1.25in}{*}{*}
\checkandfixthelayout

\pagestyle{companion}

%% Bringhurst chapter style
\makechapterstyle{bringhurst}{%
\renewcommand{\chapterheadstart}{}
\renewcommand{\printchaptername}{}
\renewcommand{\chapternamenum}{}
\renewcommand{\printchapternum}{}
\renewcommand{\afterchapternum}{}
\renewcommand{\printchaptertitle}[1]{%
\raggedright\Large\scshape\MakeLowercase{##1}}
\renewcommand{\afterchaptertitle}{%
\vskip\onelineskip \hrule\vskip\onelineskip}
}

\setsecheadstyle{\raggedright\scshape\MakeLowercase}
\setbeforesecskip{-\onelineskip}
\setaftersecskip{\onelineskip}

\chapterstyle{bringhurst}

\begin{document}
\chapter[Jude]{The General Epistle of Jude}
Jude, the servant of Jesus Christ, and brother of James, to them that
are sanctified by God the Father, and preserved in Jesus Christ,
\emph{and} called: Mercy unto you, and peace, and love, be multiplied.

Beloved, when I gave all diligence to write unto you of the common
salvation, it was needful for me to write unto you, and exhort
\emph{you} that ye should earnestly contend for the faith which was once
delivered unto the saints. For there are certain men crept in unawares,
who were before of old ordained to this condemnation, ungodly men,
turning the grace of our God into lasciviousness, and denying the only
Lord God, and our Lord Jesus Christ. I will therefore put you in
remembrance, though ye once knew this, how that the Lord, having saved
the people out of the land of Egypt, afterward destroyed them that
believed not. And the angels which kept not their first estate, but left
their own habitation, he hath reserved in everlasting chains under
darkness unto the judgment of the great day. Even as Sodom and Gomorrha,
and the cities about them in like manner, giving themselves over to
fornication, and going after strange flesh, are set forth for an
example, suffering the vengeance of eternal fire. Likewise also these
\emph{filthy} dreamers defile the flesh, despise dominion, and speak
evil of dignities. Yet Michael the archangel, when contending with the
devil he disputed about the body of Moses, durst not bring against him a
railing accusation, but said, The Lord rebuke thee. But these speak evil
of those things which they know not: but what they know naturally, as
brute beasts, in those things they corrupt themselves. Woe unto them!
for they have gone in the way of Cain, and ran greedily after the error
of Balaam for reward, and perished in the gainsaying of Core. These are
spots in your feasts of charity, when they feast with you, feeding
themselves without fear: clouds \emph{they are} without water, carried
about of winds; trees whose fruit withereth, without fruit, twice dead,
plucked up by the roots; Raging waves of the sea, foaming out their own
shame; wandering stars, to whom is reserved the blackness of darkness
for ever. And Enoch also, the seventh from Adam, prophesied of these,
saying, Behold, the Lord cometh with ten thousands of his saints, To
execute judgment upon all, and to convince all that are ungodly among
them of all their ungodly deeds which they have ungodly committed, and
of all their hard \emph{speeches} which ungodly sinners have spoken
against him. These are murmurers, complainers, walking after their own
lusts; and their mouth speaketh great swelling \emph{words,} having
men's persons in admiration because of advantage.

But, beloved, remember ye the words which were spoken before of the
apostles of our Lord Jesus Christ; How that they told you there should
be mockers in the last time, who should walk after their own ungodly
lusts. These be they who separate themselves, sensual, having not the
Spirit. But ye, beloved, building up yourselves on your most holy faith,
praying in the Holy Ghost, Keep yourselves in the love of God, looking
for the mercy of our Lord Jesus Christ unto eternal life. And of some
have compassion, making a difference: And others save with fear, pulling
\emph{them} out of the fire; hating even the garment spotted by the
flesh.

Now unto him that is able to keep you from falling, and to present
\emph{you} faultless before the presence of his glory with exceeding
joy, To the only wise God our Saviour, \emph{be} glory and majesty,
dominion and power, both now and ever. Amen.
\end{document}